---
title: '`r params$target_sector`: Technology Adoption and Productivity Analysis'
subtitle: "Vietnam Enterprise Survey: 34 CGE Sectors, 2019-2023 | Real 2019 VND"
output:
  html_document:
    theme: flatly
    toc: true
    toc_depth: 2
    df_print: paged
  pdf_document:
    toc: false
    toc_depth: '2'
params:
  target_sector: "Education"
  data_path: 'dn18_23_Cleaning&SectorAggregation_FINAL.rds'
  decomp_path: 'LP_decomposition_CGEsector.xlsx'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo      = FALSE,
  warning   = FALSE,
  message   = FALSE,
  fig.align = "center",
  fig.width = 11,
  fig.height = 6.5
)

library(tidyverse)
library(scales)
library(glue)
library(knitr)
library(kableExtra)
library(sf)
library(geodata)
library(stringi)
library(purrr)
library(splines)
library(quantreg)
library(readxl)

sf::sf_use_s2(FALSE)

base_year      <- 2019
snapshot_year  <- 2023
target_sector  <- params$target_sector
analysis_years <- 2019:2023

# ── Data loading ──────────────────────────────────────────────────────────────
# The template looks for an object called df_final in the R environment first.
# If not found, it reads from params$data_path. The default relative path
# "dn18_23_Cleaning&SectorAggregation_FINAL.rds" assumes the .rds file sits in
# the same folder as this .Rmd. Adjust params$data_path if it is elsewhere.
if (!exists("df_final")) {
  if (!is.null(params$data_path) && params$data_path != "") {
    df_final <- readr::read_rds(params$data_path)
  } else {
    stop(
      "df_final is not loaded. Supply params$data_path pointing to ",
      "dn18_23_Cleaning&SectorAggregation_FINAL.rds, or load df_final ",
      "into your R session before rendering."
    )
  }
}

# ── Helper functions ──────────────────────────────────────────────────────────

safe_divide <- function(numerator, denominator) {
  ifelse(is.na(denominator) | denominator == 0, NA_real_, numerator / denominator)
}

# Returns a colour-coded HTML span with a directional arrow and percentage.
format_change <- function(x) {
  arrow  <- ifelse(x >= 0, "↑", "↓")
  colour <- ifelse(x >= 0, "#15803D", "#DC2626")
  sign   <- ifelse(x >= 0, "+", "")
  paste0(
    "<span style='color:", colour, "; font-weight:700;'>",
    arrow, " ", sign, percent(x, accuracy = 0.1),
    "</span>"
  )
}

# Returns a plain-English description of growth magnitude.
growth_strength <- function(x) {
  case_when(
    is.na(x)   ~ "showed limited measurable change",
    x >= 0.20  ~ "increased strongly",
    x >= 0.05  ~ "increased moderately",
    x > -0.05  ~ "remained broadly stable",
    x > -0.20  ~ "declined moderately",
    TRUE        ~ "declined strongly"
  )
}

# Strips accents and non-alphanumeric characters for province name matching.
clean_name <- function(x) {
  x %>%
    stringi::stri_trans_general("Latin-ASCII") %>%
    str_to_lower() %>%
    str_replace_all("[^a-z0-9]", "")
}

# Wraps interpretation text in a styled left-bordered div for consistent
# visual formatting across all sections.
interp <- function(...) {
  text <- paste0(...)
  text <- gsub("\\*\\*(.*?)\\*\\*", "<strong>\\1</strong>", text, perl = TRUE)
  cat(paste0('<div class="interpretation">\n\n', text, '\n\n</div>\n'))
}

theme_sector <- function() {
  theme_minimal(base_size = 12) +
    theme(
      plot.title       = element_text(face = "bold", size = 16, colour = "#12355B"),
      plot.subtitle    = element_text(size = 10.5, colour = "#5B6F8A", lineheight = 1.25),
      axis.title       = element_text(face = "bold"),
      legend.position  = "bottom",
      panel.grid.minor = element_blank(),
      plot.margin      = margin(10, 15, 10, 15)
    )
}
```

```{r css, results='asis', echo=FALSE}
cat("
<style>

/* ── Google Font (Inter) ─────────────────────────────────────────────────── */
@import url('https://fonts.googleapis.com/css2?family=Inter:ital,wght@0,300;0,400;0,600;0,700;1,400&display=swap');

/* ── Base typography ─────────────────────────────────────────────────────── */
body {
  font-family: 'Inter', sans-serif;
  font-size: 15px;
  line-height: 1.70;
  color: #1F2933;
}

/* ── Headings ────────────────────────────────────────────────────────────── */
h1 {
  color: #12355B;
  font-weight: 700;
  border-bottom: 3px solid #2A9D8F;
  padding-bottom: 10px;
  margin-top: 48px;
  margin-bottom: 6px;
}
h2 {
  color: #12355B;
  font-weight: 600;
  font-size: 1.15em;
  border-left: 4px solid #2A9D8F;
  padding-left: 12px;
  margin-top: 32px;
  margin-bottom: 4px;
}
h3 { color: #12355B; font-weight: 600; }

/* ── One-line section description below each heading ─────────────────────── */
.section-desc {
  color: #5B6F8A;
  font-size: 13.5px;
  margin: -4px 0 16px 0;
  font-style: italic;
}

/* ── Interpretation block ────────────────────────────────────────────────── */
.interpretation {
  background: #EDF4FB;
  border-left: 4px solid #2A9D8F;
  padding: 14px 18px 12px 18px;
  border-radius: 0 6px 6px 0;
  margin: 12px 0 24px 0;
  font-size: 14px;
  line-height: 1.75;
}
.interpretation::before {
  content: 'ANALYTICAL NOTE';
  display: block;
  font-size: 10px;
  font-weight: 700;
  letter-spacing: 1.5px;
  color: #2A9D8F;
  margin-bottom: 8px;
}

/* ── Data quality callout ────────────────────────────────────────────────── */
.data-note {
  background: #FFFBF0;
  border-left: 4px solid #F4A261;
  padding: 10px 14px;
  border-radius: 0 6px 6px 0;
  margin: 8px 0 16px 0;
  font-size: 13px;
  color: #334E68;
}

/* ── Readme box ──────────────────────────────────────────────────────────── */
.readme-box {
  background: #F8F9FA;
  border: 1px solid #D0DAE6;
  border-radius: 8px;
  padding: 18px 22px;
  margin-bottom: 28px;
  font-size: 14px;
}
.readme-box pre { background: #E8ECEF; padding: 10px 14px; border-radius: 4px; font-size: 13px; }
.readme-box code { font-size: 13px; }

/* ── Misc ────────────────────────────────────────────────────────────────── */
.caption { color: #6B7280; font-size: 12px; }
.chart-card {
  padding: 16px 18px;
  border: 1px solid #E5E7EB;
  border-radius: 12px;
  margin-bottom: 22px;
  background: #FFFFFF;
}

/* ── Cover block ─────────────────────────────────────────────────────────── */
.cover-block {
  background: linear-gradient(135deg, #12355B 0%, #1a4a7a 100%);
  color: #FFFFFF;
  padding: 36px 40px 32px 40px;
  border-radius: 10px;
  margin-bottom: 36px;
}
.cover-block .sector-name {
  font-size: 26px;
  font-weight: 700;
  line-height: 1.3;
  margin-bottom: 6px;
}
.cover-block .cover-subtitle {
  font-size: 14px;
  color: #A8C7E8;
  margin-bottom: 20px;
  letter-spacing: 0.3px;
}
.cover-block .cover-meta {
  display: flex;
  gap: 32px;
  flex-wrap: wrap;
  border-top: 1px solid rgba(255,255,255,0.2);
  padding-top: 16px;
}
.cover-block .meta-item {
  font-size: 12px;
  color: #C8DCF0;
}
.cover-block .meta-item strong {
  display: block;
  color: #FFFFFF;
  font-size: 13px;
  margin-bottom: 2px;
}

/* ── Part divider ────────────────────────────────────────────────────────── */
.part-divider {
  border-top: 2px solid #2A9D8F;
  margin: 52px 0 36px 0;
  text-align: center;
  position: relative;
}
.part-divider span {
  background: white;
  padding: 0 18px;
  color: #5B6F8A;
  font-size: 11px;
  font-weight: 600;
  letter-spacing: 1.8px;
  text-transform: uppercase;
  position: relative;
  top: -10px;
  font-family: 'Inter', sans-serif;
}

/* ── Figure captions ─────────────────────────────────────────────────────── */
.figure .caption, caption {
  font-size: 12.5px;
  color: #5B6F8A;
  font-style: italic;
  text-align: left;
  margin-top: 6px;
  line-height: 1.5;
}


/* ── Section 4 insight blocks ─────────────────────────────────────────────── */
.highlight-grid{display:grid;grid-template-columns:repeat(auto-fit,minmax(260px,1fr));gap:14px;margin:16px 0 24px 0;}
.highlight-block{border:1px solid #D9E1E8;border-top:4px solid #2A9D8F;border-radius:8px;padding:14px 16px;background:#fff;box-shadow:0 1px 4px rgba(18,53,91,.06);}
.highlight-title{display:block;font-weight:700;color:#12355B;margin-bottom:6px;font-size:14px;}
.highlight-body{color:#334E68;font-size:13.5px;line-height:1.65;}
.policy-box{background:#F8FAFC;border-left:4px solid #12355B;border-radius:0 8px 8px 0;padding:16px 18px;margin:14px 0 20px 0;font-size:14px;line-height:1.75;}
.policy-box strong{color:#12355B;}
.strategy-close{background:linear-gradient(135deg,#F1F7F9 0%,#fff 100%);border:1px solid #D9E1E8;border-radius:10px;padding:18px 22px;margin:16px 0 28px 0;color:#12355B;font-size:15px;line-height:1.75;}

/* ── Print styles ────────────────────────────────────────────────────────── */
@media print {
  .cover-block { -webkit-print-color-adjust: exact; print-color-adjust: exact; }
  body { font-size: 12px; line-height: 1.55; }
  h1 { font-size: 18px; margin-top: 24px; }
  h2 { font-size: 14px; }
  .interpretation { break-inside: avoid; }
  img { max-width: 100%; page-break-inside: avoid; }
  .part-divider { margin: 24px 0 18px 0; }
}

</style>
")
```

```{=html}
<!--
HOW TO RUN THIS REPORT
======================

Dataset:
  Place "dn18_23_Cleaning&SectorAggregation_FINAL.rds" in the same folder as
  this .Rmd, or update params$data_path to the correct absolute path.

Required R packages:
  tidyverse, scales, glue, knitr, kableExtra, sf, geodata, stringi, purrr

Option 1 - Knit interactively in RStudio:
  Click Knit -> Knit with Parameters, type the sector name in the dialog,
  and click Knit. The sector name must exactly match a CGE_sectorname value
  in the dataset.

Option 2 - Render from the R console:
  rmarkdown::render(
    input         = "sector_report_template_complete.Rmd",
    params        = list(target_sector = "ICT and publishing"),
    output_file   = "output/ICT_and_publishing_report.html",
    knit_root_dir = getwd()
  )

Note: A batch render script for all 34 CGE sectors will be provided separately
after this template is validated on individual sectors.
-->
```

```{r data-preparation, include=FALSE}
# -------------------------------------------------------------------
# Filter to target sector
# -------------------------------------------------------------------

sector_df <- df_final %>%
  filter(CGE_sectorname == target_sector)

if (nrow(sector_df) == 0) {
  stop(glue(
    "No data found for target_sector = '{target_sector}'. ",
    "Check that the name exactly matches a CGE_sectorname value in the dataset."
  ))
}

df_page1 <- df_final

# -------------------------------------------------------------------
# sector_year: one row per survey year for the target sector.
# All monetary values are already in real million VND (2019 base).
# Intensity metrics use ratio-of-totals (sum numerator / sum denominator)
# to avoid giving equal weight to large and small firms.
# -------------------------------------------------------------------

sector_year <- sector_df %>%
  group_by(survey_year) %>%
  summarise(
    firms                              = n_distinct(tax_id),
    labour                             = sum(labour_headcount_end, na.rm = TRUE),
    female_labour                      = sum(labour_female_headcount_end, na.rm = TRUE),
    revenue                            = sum(total_net_revenue, na.rm = TRUE),
    capital                            = sum(capital_sources_total_end, na.rm = TRUE),
    total_investment                   = sum(total_investment, na.rm = TRUE),
    equipment_investment               = sum(investment_machinery_equipment, na.rm = TRUE),
    innovation_investment              = sum(innovation_investment_proxy, na.rm = TRUE),
    firms_with_any_selected_investment = n_distinct(tax_id[has_equipment_investment == 1 | has_innovation_investment == 1]),
    equipment_adoption_rate            = safe_divide(n_distinct(tax_id[has_equipment_investment == 1]), n_distinct(tax_id)),
    innovation_adoption_rate           = safe_divide(n_distinct(tax_id[has_innovation_investment == 1]), n_distinct(tax_id)),
    female_labour_share                = safe_divide(female_labour, labour),
    labour_productivity                = safe_divide(revenue, labour),
    .groups = "drop"
  )

# all_sector_year: same structure but for every CGE sector, used for benchmarks.
all_sector_year <- df_final %>%
  group_by(CGE_sectorname, survey_year) %>%
  summarise(
    firms               = n_distinct(tax_id),
    labour              = sum(labour_headcount_end, na.rm = TRUE),
    revenue             = sum(total_net_revenue, na.rm = TRUE),
    labour_productivity = safe_divide(revenue, labour),
    .groups = "drop"
  )

# -------------------------------------------------------------------
# Firm dynamics: entry, exit, and staying counts.
# Used in the Page 1 KPI table and Visual 3.
# -------------------------------------------------------------------

firm_year_base <- sector_df %>%
  filter(!is.na(tax_id), !is.na(survey_year)) %>%
  distinct(tax_id, survey_year)

all_years <- sort(unique(firm_year_base$survey_year))

firm_flow_data <- map_dfr(all_years, function(y) {
  current_firms  <- firm_year_base %>% filter(survey_year == y)     %>% pull(tax_id) %>% unique()
  previous_firms <- firm_year_base %>% filter(survey_year == y - 1) %>% pull(tax_id) %>% unique()
  tibble(
    survey_year        = y,
    staying            = ifelse(y == min(all_years), length(current_firms), length(intersect(current_firms, previous_firms))),
    entry              = ifelse(y == min(all_years), 0, length(setdiff(current_firms, previous_firms))),
    exit               = ifelse(y == min(all_years), 0, length(setdiff(previous_firms, current_firms))),
    total_active_firms = length(current_firms)
  )
})

output_index_data <- sector_df %>%
  group_by(survey_year) %>%
  summarise(total_output = sum(total_net_revenue, na.rm = TRUE), .groups = "drop") %>%
  arrange(survey_year) %>%
  mutate(
    output_base  = total_output[survey_year == base_year][1],
    output_index = safe_divide(total_output, output_base) * 100
  ) %>%
  select(survey_year, output_index)

firm_flow_data <- firm_flow_data %>%
  left_join(output_index_data, by = "survey_year") %>%
  arrange(survey_year)

kpi_firm_flow_snapshot <- firm_flow_data %>%
  filter(survey_year == snapshot_year) %>%
  transmute(
    staying_firms           = staying,
    entry_firms             = entry,
    exit_firms              = exit,
    staying_rate            = safe_divide(staying, total_active_firms),
    entry_rate              = safe_divide(entry, total_active_firms),
    exit_rate               = safe_divide(exit, staying + exit),
    sector_switch_out_firms = 0,
    sector_switch_out_rate  = 0
  )

bar_positive <- firm_flow_data %>%
  select(survey_year, staying, entry) %>%
  pivot_longer(cols = c(staying, entry), names_to = "Firm flow", values_to = "Number of firms") %>%
  mutate(
    `Firm flow` = recode(`Firm flow`, staying = "Staying (incumbents)", entry = "Entry"),
    x_pos = as.numeric(factor(survey_year)) - 0.16
  )

bar_exit <- firm_flow_data %>%
  transmute(
    survey_year,
    `Firm flow`       = "Exit",
    `Number of firms` = -exit,
    x_pos             = as.numeric(factor(survey_year)) + 0.22
  )

firm_axis_max   <- max(firm_flow_data$staying + firm_flow_data$entry, na.rm = TRUE)
firm_axis_min   <- -max(firm_flow_data$exit, na.rm = TRUE)
index_min       <- min(firm_flow_data$output_index, na.rm = TRUE)
index_max       <- max(firm_flow_data$output_index, na.rm = TRUE)
index_padding   <- (index_max - index_min) * 0.10
index_min_plot  <- index_min - index_padding
index_max_plot  <- index_max + index_padding
scale_factor_v3 <- ifelse(
  index_max_plot > index_min_plot,
  (firm_axis_max - firm_axis_min) / (index_max_plot - index_min_plot),
  1
)
```

::: part-divider
<span>Part I: Sector Structure and Performance</span>
:::

# 1. Sector Structure and Performance

## 1.1 KPI Snapshot

<p class="section-desc">

Key performance indicators for `r snapshot_year` compared with the `r base_year` baseline, covering firm count, employment, revenue scale, capital intensity, and firm entry/exit dynamics.

</p>

```{r page1-kpi, results='asis'}
kpi_base <- sector_year %>%
  filter(survey_year == base_year) %>%
  summarise(
    firms_base                 = sum(firms, na.rm = TRUE),
    labour_base                = sum(labour, na.rm = TRUE),
    revenue_base               = sum(revenue, na.rm = TRUE),
    female_labour_share_base   = safe_divide(sum(female_labour, na.rm = TRUE), sum(labour, na.rm = TRUE)),
    revenue_per_firm_base      = safe_divide(sum(revenue, na.rm = TRUE), sum(firms, na.rm = TRUE)),
    capital_per_worker_base    = safe_divide(sum(capital, na.rm = TRUE), sum(labour, na.rm = TRUE)),
    .groups = "drop"
  )

kpi_snapshot_raw <- sector_year %>%
  filter(survey_year == snapshot_year) %>%
  summarise(
    firms                = sum(firms, na.rm = TRUE),
    labour               = sum(labour, na.rm = TRUE),
    revenue              = sum(revenue, na.rm = TRUE),
    female_labour_share  = safe_divide(sum(female_labour, na.rm = TRUE), sum(labour, na.rm = TRUE)),
    revenue_per_firm     = safe_divide(sum(revenue, na.rm = TRUE), sum(firms, na.rm = TRUE)),
    capital_per_worker   = safe_divide(sum(capital, na.rm = TRUE), sum(labour, na.rm = TRUE)),
    .groups = "drop"
  ) %>%
  bind_cols(kpi_base) %>%
  bind_cols(kpi_firm_flow_snapshot) %>%
  mutate(
    firms_growth               = safe_divide(firms, firms_base) - 1,
    labour_growth              = safe_divide(labour, labour_base) - 1,
    revenue_growth             = safe_divide(revenue, revenue_base) - 1,
    female_labour_share_change = female_labour_share - female_labour_share_base,
    revenue_per_firm_growth    = safe_divide(revenue_per_firm, revenue_per_firm_base) - 1,
    capital_per_worker_growth  = safe_divide(capital_per_worker, capital_per_worker_base) - 1
  )

kpi_table <- tibble(
  `Left KPI` = c("Active firms", "Total labour", "Average revenue per firm", "Firm dynamics"),
  `Left Value` = c(
    paste0(comma(kpi_snapshot_raw$firms), " firms (", format_change(kpi_snapshot_raw$firms_growth), " vs 2019)"),
    paste0(comma(kpi_snapshot_raw$labour), " workers (", format_change(kpi_snapshot_raw$labour_growth), " vs 2019)"),
    paste0(comma(round(kpi_snapshot_raw$revenue_per_firm, 1)), " mn VND/firm (", format_change(kpi_snapshot_raw$revenue_per_firm_growth), " vs 2019)"),
    paste0("Stay: ", comma(kpi_snapshot_raw$staying_firms), " (", percent(kpi_snapshot_raw$staying_rate, accuracy = 0.1), ")", " | Entry: ", comma(kpi_snapshot_raw$entry_firms), " (", percent(kpi_snapshot_raw$entry_rate, accuracy = 0.1), ")")
  ),
  `Right KPI` = c("Total revenue", "Female labour share", "Capital Intensity", "Firm dynamics"),
  `Right Value` = c(
    paste0(comma(round(kpi_snapshot_raw$revenue / 1e6, 2)), " trillion VND (", format_change(kpi_snapshot_raw$revenue_growth), " vs 2019)"),
    paste0(percent(kpi_snapshot_raw$female_labour_share, accuracy = 0.1), " (", format_change(kpi_snapshot_raw$female_labour_share_change), " percentage points vs 2019)"),
    paste0(comma(round(kpi_snapshot_raw$capital_per_worker, 1)), " mn VND/worker (", format_change(kpi_snapshot_raw$capital_per_worker_growth), " vs 2019)"),
    paste0("Exit: ", comma(kpi_snapshot_raw$exit_firms), " (", percent(kpi_snapshot_raw$exit_rate, accuracy = 0.1), ")", " | Switch-out: ", comma(kpi_snapshot_raw$sector_switch_out_firms), " (", percent(kpi_snapshot_raw$sector_switch_out_rate, accuracy = 0.1), ")")
  )
)

kpi_table %>%
  kable(
    format = "html", escape = FALSE,
    caption = glue("{target_sector}: Page 1 KPI Snapshot, {snapshot_year}"),
    col.names = c("", "", "", "")
  ) %>%
  kable_styling(
    full_width = TRUE,
    bootstrap_options = c("striped", "hover", "condensed"),
    font_size = 14
  ) %>%
  column_spec(1, bold = TRUE, color = "#12355B", width = "20%", extra_css = "white-space: nowrap;") %>%
  column_spec(2, color = "#1F2933", width = "28%") %>%
  column_spec(3, bold = TRUE, color = "#12355B", width = "20%", extra_css = "white-space: nowrap;") %>%
  column_spec(4, color = "#1F2933", width = "32%") %>%
  row_spec(4, extra_css = "border-top: 2px solid #D9E1E8;")
```

```{r page1-kpi-interpretation, results='asis'}
net_firm_change <- kpi_snapshot_raw$entry_firms - kpi_snapshot_raw$exit_firms
net_direction   <- if_else(net_firm_change >= 0, "net gain", "net loss")
scale_msg       <- if_else(kpi_snapshot_raw$revenue_per_firm_growth >= 0, "improving", "deteriorating")

interp(glue(
  "**{snapshot_year} snapshot:** {target_sector} comprised **{comma(kpi_snapshot_raw$firms)}** active firms generating ",
  "**{comma(round(kpi_snapshot_raw$revenue / 1e6, 2))} trillion VND** in real revenue ",
  "({format_change(kpi_snapshot_raw$revenue_growth)} vs {base_year}), employing **{comma(kpi_snapshot_raw$labour)}** workers ",
  "({format_change(kpi_snapshot_raw$labour_growth)} vs {base_year}). Average revenue per firm stood at ",
  "**{comma(round(kpi_snapshot_raw$revenue_per_firm, 1))} million VND** ({format_change(kpi_snapshot_raw$revenue_per_firm_growth)} vs {base_year}), ",
  "indicating {scale_msg} firm-level scale. ",
  "**Firm dynamics:** **{comma(kpi_snapshot_raw$entry_firms)}** firms entered and **{comma(kpi_snapshot_raw$exit_firms)}** exited ",
  "in {snapshot_year}, a {net_direction} of **{comma(abs(net_firm_change))}** firms. ",
  "**{comma(kpi_snapshot_raw$staying_firms)}** firms ({percent(kpi_snapshot_raw$staying_rate, accuracy = 0.1)}) ",
  "were continuing incumbents."
))
```

## 1.2 Revenue Scale and Labour Benchmark

<p class="section-desc">

Sector total revenue and labour headcount compared against the all-sector average, `r base_year`–`r snapshot_year`. Left axis: revenue (trillion VND). Right axis: labour headcount.

</p>

```{r page1-visual1-prep, include=FALSE}
labour_selected_label  <- glue("{target_sector} Labour")
labour_average_label   <- "Average Labour - All Sectors"
selected_revenue_label <- glue("{target_sector} Sector Revenue")
average_revenue_label  <- "Average Revenue - All Sectors"

all_sector_benchmark <- all_sector_year %>%
  group_by(survey_year) %>%
  summarise(
    mean_sector_revenue = mean(revenue, na.rm = TRUE),
    mean_sector_labour  = mean(labour, na.rm = TRUE),
    .groups = "drop"
  )

v1_sector <- sector_year %>%
  left_join(all_sector_benchmark, by = "survey_year") %>%
  arrange(survey_year) %>%
  transmute(
    survey_year,
    !!selected_revenue_label := revenue,
    !!average_revenue_label  := mean_sector_revenue,
    !!labour_selected_label  := labour,
    !!labour_average_label   := mean_sector_labour
  )

v1_bar <- v1_sector %>%
  pivot_longer(
    cols = c(all_of(selected_revenue_label), all_of(average_revenue_label)),
    names_to = "Revenue benchmark", values_to = "Revenue"
  ) %>%
  mutate(
    revenue_trillion    = Revenue / 1e6,
    `Revenue benchmark` = factor(`Revenue benchmark`, levels = c(selected_revenue_label, average_revenue_label))
  )

v1_line <- v1_sector %>%
  pivot_longer(
    cols = c(all_of(labour_selected_label), all_of(labour_average_label)),
    names_to = "Labour benchmark", values_to = "Labour"
  ) %>%
  mutate(
    `Labour benchmark` = factor(`Labour benchmark`, levels = c(labour_selected_label, labour_average_label))
  )

max_revenue    <- max(v1_bar$revenue_trillion, na.rm = TRUE)
max_labour     <- max(v1_line$Labour, na.rm = TRUE)
scale_factor_v1 <- ifelse(max_labour > 0, max_revenue / max_labour, 1)
```

```{r page1-visual1, fig.cap=paste0('Figure 1.1: Total sector revenue and labour headcount benchmarked against the all-sector average, ', base_year, '–', snapshot_year, '.')}
ggplot() +
  geom_col(
    data = v1_bar,
    aes(x = factor(survey_year), y = revenue_trillion, fill = `Revenue benchmark`),
    position = position_dodge(width = 0.7), width = 0.6, colour = "white", linewidth = 0.35
  ) +
  geom_line(
    data = v1_line,
    aes(x = factor(survey_year), y = Labour * scale_factor_v1, colour = `Labour benchmark`, group = `Labour benchmark`),
    linewidth = 1.25
  ) +
  geom_point(
    data = v1_line,
    aes(x = factor(survey_year), y = Labour * scale_factor_v1, colour = `Labour benchmark`, shape = `Labour benchmark`),
    size = 3
  ) +
  geom_text(
    data = v1_bar,
    aes(x = factor(survey_year), y = revenue_trillion, label = comma(round(revenue_trillion, 2)),
        group = `Revenue benchmark`),
    position = position_dodge(width = 0.7),
    vjust = -0.45, size = 3, fontface = "bold", colour = "#12355B"
  ) +
  scale_y_continuous(
    name   = "Revenue (trillion VND)",
    labels = comma_format(accuracy = 0.01),
    expand = expansion(mult = c(0, 0.15)),
    sec.axis = sec_axis(~ . / scale_factor_v1, name = "Labour headcount", labels = comma_format(accuracy = 1))
  ) +
  scale_fill_manual(
    values = setNames(c("#2A9D8F", "#D9E1E8"), c(selected_revenue_label, average_revenue_label)),
    name = "Revenue benchmark", drop = FALSE
  ) +
  scale_colour_manual(
    values = setNames(c("#E76F51", "#6B7280"), c(labour_selected_label, labour_average_label)),
    name = "Labour benchmark", drop = FALSE
  ) +
  scale_shape_manual(
    values = setNames(c(16, 15), c(labour_selected_label, labour_average_label)),
    name = "Labour benchmark"
  ) +
  labs(
    title    = glue("{target_sector}: Revenue Scale and Labour Benchmark"),
    subtitle = paste(
      glue("Bars compare {target_sector} revenue with average all-sector revenue; lines compare total labour headcount"),
      "\nLeft axis measures revenue (trillion VND). Right axis measures labour headcount."
    ),
    x = NULL
  ) +
  guides(
    fill   = guide_legend(order = 1),
    colour = guide_legend(order = 2, override.aes = list(linewidth = 1.3, size = 3)),
    shape  = guide_legend(order = 2)
  ) +
  theme_sector() +
  theme(
    legend.position      = "bottom",
    legend.box           = "vertical",
    legend.title         = element_text(face = "bold"),
    axis.title.y.left    = element_text(face = "bold", colour = "#2A9D8F"),
    axis.text.y.left     = element_text(colour = "#2A9D8F"),
    axis.title.y.right   = element_text(face = "bold", colour = "#E76F51"),
    axis.text.y.right    = element_text(colour = "#E76F51")
  )
```

```{r page1-visual1-interpretation, results='asis'}
v1_stats <- sector_year %>%
  filter(survey_year %in% c(base_year, snapshot_year)) %>%
  summarise(
    revenue_growth = safe_divide(revenue[survey_year == snapshot_year], revenue[survey_year == base_year]) - 1,
    labour_growth  = safe_divide(labour[survey_year == snapshot_year], labour[survey_year == base_year]) - 1
  )

rev_2019        <- sector_year %>% filter(survey_year == base_year) %>% pull(revenue)
rev_2023        <- sector_year %>% filter(survey_year == snapshot_year) %>% pull(revenue)
benchmark_2019  <- all_sector_benchmark %>% filter(survey_year == base_year) %>% pull(mean_sector_revenue)
benchmark_2023  <- all_sector_benchmark %>% filter(survey_year == snapshot_year) %>% pull(mean_sector_revenue)
rev_ratio       <- safe_divide(rev_2023, benchmark_2023)
rev_ratio_base  <- safe_divide(rev_2019, benchmark_2019)
rev_ratio_msg   <- if_else(
  rev_ratio >= 1,
  glue("{round(rev_ratio, 1)}x the all-sector average"),
  glue("{percent(rev_ratio, accuracy = 1)} of the all-sector average")
)

benchmark_gap_msg <- case_when(
  is.na(rev_ratio_base) | is.na(rev_ratio) ~ "",
  rev_ratio > rev_ratio_base + 0.2 ~ glue(
    "The sector's revenue advantage over the all-sector average widened ",
    "(from {round(rev_ratio_base, 1)}x in {base_year} to {round(rev_ratio, 1)}x in {snapshot_year}), ",
    "indicating that the sector expanded its relative economic footprint."
  ),
  rev_ratio < rev_ratio_base - 0.2 ~ glue(
    "The sector's revenue position relative to the all-sector average narrowed ",
    "(from {round(rev_ratio_base, 1)}x in {base_year} to {round(rev_ratio, 1)}x in {snapshot_year}), ",
    "suggesting other sectors grew faster over the period."
  ),
  TRUE ~ "The sector's revenue position relative to the all-sector average remained broadly stable across the period."
)

revenue_labour_msg <- case_when(
  v1_stats$revenue_growth > v1_stats$labour_growth + 0.05 ~
    "Revenue grew materially faster than employment, consistent with improving revenue intensity and rising labour productivity at the sector level.",
  v1_stats$labour_growth > v1_stats$revenue_growth + 0.05 ~
    "Employment expanded faster than revenue, suggesting the sector absorbed labour without equivalent productivity gains over the period.",
  TRUE ~
    "Revenue and employment grew at broadly similar rates, indicating stable revenue per worker rather than productivity improvement."
)

interp(glue(
  "Between {base_year} and {snapshot_year}, {target_sector} total revenue {growth_strength(v1_stats$revenue_growth)} ",
  "({format_change(v1_stats$revenue_growth)}), reaching **{comma(round(rev_2023 / 1e6, 2))} trillion VND**, ",
  "which is {rev_ratio_msg} in {snapshot_year}. ",
  "{benchmark_gap_msg} ",
  "Employment {growth_strength(v1_stats$labour_growth)} ({format_change(v1_stats$labour_growth)}) over the same period. ",
  "{revenue_labour_msg}"
))
```

## 1.3 Five-Year Revenue Share by Ownership Type

<p class="section-desc">

Cumulative revenue share by ownership type (SOE, domestic private, FDI) pooled across `r base_year`–`r snapshot_year`, benchmarked against the all-sector economy.

</p>

```{r page1-visual2-prep, include=FALSE}
ownership_palette_benchmark <- c("Selected sector" = "#2A9D8F", "All sectors" = "#D9E1E8")
ownership_levels <- c("SOE", "Domestic private firms", "FDI")

sector_ownership_share_5yr <- sector_df %>%
  filter(survey_year %in% analysis_years, !is.na(firm_ownership_type_factor)) %>%
  mutate(Ownership = factor(as.character(firm_ownership_type_factor), levels = ownership_levels)) %>%
  filter(!is.na(Ownership)) %>%
  group_by(Ownership) %>%
  summarise(revenue = sum(total_net_revenue, na.rm = TRUE), .groups = "drop") %>%
  tidyr::complete(Ownership = factor(ownership_levels, levels = ownership_levels), fill = list(revenue = 0)) %>%
  mutate(revenue_share = revenue / sum(revenue, na.rm = TRUE), Benchmark = "Selected sector")

all_sector_ownership_share_5yr <- df_final %>%
  filter(survey_year %in% analysis_years, !is.na(firm_ownership_type_factor)) %>%
  mutate(Ownership = factor(as.character(firm_ownership_type_factor), levels = ownership_levels)) %>%
  filter(!is.na(Ownership)) %>%
  group_by(Ownership) %>%
  summarise(revenue = sum(total_net_revenue, na.rm = TRUE), .groups = "drop") %>%
  tidyr::complete(Ownership = factor(ownership_levels, levels = ownership_levels), fill = list(revenue = 0)) %>%
  mutate(revenue_share = revenue / sum(revenue, na.rm = TRUE), Benchmark = "All sectors")

ownership_grouped_data <- bind_rows(sector_ownership_share_5yr, all_sector_ownership_share_5yr) %>%
  mutate(
    Ownership = factor(Ownership, levels = ownership_levels),
    Benchmark = factor(Benchmark, levels = c("Selected sector", "All sectors")),
    bar_label = percent(revenue_share, accuracy = 0.1)
  )
```

```{r page1-visual2, fig.cap=paste0('Figure 1.2: Five-year cumulative revenue share by ownership type (SOE, domestic private, FDI), ', base_year, '–', snapshot_year, ', benchmarked against all sectors.')}
ggplot(ownership_grouped_data, aes(x = Ownership, y = revenue_share, fill = Benchmark)) +
  geom_col(position = position_dodge(width = 0.72), width = 0.62, colour = "white", linewidth = 0.35) +
  geom_text(
    aes(label = bar_label),
    position = position_dodge(width = 0.72), vjust = -0.35,
    size = 3.8, fontface = "bold", colour = "#12355B"
  ) +
  scale_y_continuous(
    labels = percent_format(accuracy = 1),
    limits = c(0, max(ownership_grouped_data$revenue_share, na.rm = TRUE) * 1.18),
    expand = expansion(mult = c(0, 0.02))
  ) +
  scale_fill_manual(values = ownership_palette_benchmark, name = "Benchmark") +
  labs(
    title    = glue("{target_sector}: Five-Year Revenue Share by Ownership Type"),
    subtitle = "Revenue share by ownership type across 2019–2023, compared with all sectors",
    x = NULL, y = "Revenue share"
  ) +
  theme_sector() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(colour = "#E6EAF0", linewidth = 0.35),
    axis.text.x        = element_text(face = "bold", colour = "#12355B", size = 11),
    axis.text.y        = element_text(colour = "#5B6F8A", size = 10),
    legend.title       = element_text(face = "bold"),
    legend.text        = element_text(size = 10, colour = "#5B6F8A"),
    plot.margin        = margin(10, 20, 10, 20)
  )
```

```{r page1-visual2-interpretation, results='asis'}
ownership_ranked <- ownership_grouped_data %>%
  filter(Benchmark == "Selected sector") %>%
  arrange(desc(revenue_share))

dominant_owner <- ownership_ranked %>% slice(1)
second_owner   <- ownership_ranked %>% slice(2)

all_dominant <- ownership_grouped_data %>%
  filter(Benchmark == "All sectors", Ownership == dominant_owner$Ownership) %>%
  pull(revenue_share)
all_second <- ownership_grouped_data %>%
  filter(Benchmark == "All sectors", Ownership == second_owner$Ownership) %>%
  pull(revenue_share)

gap_dominant <- dominant_owner$revenue_share - all_dominant
gap_second   <- second_owner$revenue_share - all_second

gap_dominant_msg <- if_else(
  gap_dominant > 0,
  glue("{percent(gap_dominant, accuracy = 0.1)} percentage points above the all-sector benchmark"),
  glue("{percent(abs(gap_dominant), accuracy = 0.1)} percentage points below the all-sector benchmark")
)
gap_second_dir <- if_else(gap_second >= 0, "above", "below")

structural_ownership_msg <- case_when(
  as.character(dominant_owner$Ownership) == "FDI" & dominant_owner$revenue_share > 0.50 ~
    "FDI-dominated revenue structures typically bring higher capital intensity and technology exposure, but may limit domestic capability spillovers if supplier linkages with local firms are weak.",
  as.character(dominant_owner$Ownership) == "SOE" & dominant_owner$revenue_share > 0.50 ~
    "State-owned enterprise dominance often reflects regulated or capital-intensive activities; productivity performance in such sectors tends to depend heavily on reform of managerial incentives rather than investment volume alone.",
  as.character(dominant_owner$Ownership) == "Domestic private firms" & dominant_owner$revenue_share > 0.50 ~
    "A domestic-private-led revenue structure places firms that face competitive market pressure at the centre of the sector; productivity gains tend to be driven by selection, entry and capability upgrading rather than by state direction.",
  TRUE ~
    "The mixed ownership profile means productivity dynamics are shaped by interactions between firm types with different incentive structures, investment horizons and technology access."
)

interp(glue(
  "Pooled across {base_year}-{snapshot_year}, **{dominant_owner$Ownership}** accounts for the largest share of ",
  "{target_sector} revenue at **{percent(dominant_owner$revenue_share, accuracy = 0.1)}**, {gap_dominant_msg}. ",
  "**{second_owner$Ownership}** is the second-largest contributor at **{percent(second_owner$revenue_share, accuracy = 0.1)}**, ",
  "{gap_second_dir} its all-sector equivalent (**{percent(all_second, accuracy = 0.1)}**). ",
  "{structural_ownership_msg}"
))
```

## 1.4 Firm Entry and Exit Dynamics

<p class="section-desc">

Annual counts of entering, exiting, and staying firms alongside the sector output index (`r base_year` = 100), showing whether firm turnover is accompanied by output growth or decline.

</p>

```{r page1-visual3, fig.cap=paste0('Figure 1.3: Annual firm entry, exit, and incumbent counts with sector output index (', base_year, ' = 100), ', base_year, '–', snapshot_year, '.')}
firm_flow_data <- firm_flow_data %>%
  mutate(
    x_pos               = as.numeric(factor(survey_year)),
    output_index_scaled = (output_index - index_min_plot) * scale_factor_v3 + firm_axis_min
  )

ggplot() +
  geom_hline(yintercept = 0, colour = "#6B7280", linewidth = 0.7) +
  geom_col(
    data = bar_positive,
    aes(x = x_pos, y = `Number of firms`, fill = `Firm flow`),
    width = 0.46, colour = "white", linewidth = 0.3
  ) +
  geom_col(
    data = bar_exit,
    aes(x = x_pos, y = `Number of firms`, fill = `Firm flow`),
    width = 0.28, colour = "white", linewidth = 0.3
  ) +
  geom_smooth(
    data = firm_flow_data,
    aes(x = x_pos, y = output_index_scaled, colour = "Output index (2019 = 100)", group = 1),
    method = "loess", se = FALSE, linewidth = 1.4, span = 0.8
  ) +
  geom_point(
    data = firm_flow_data,
    aes(x = x_pos, y = output_index_scaled, colour = "Output index (2019 = 100)"),
    size = 3
  ) +
  geom_text(
    data = firm_flow_data,
    aes(x = x_pos - 0.16, y = total_active_firms, label = comma(total_active_firms)),
    vjust = -0.5, size = 3.3, fontface = "bold", colour = "#12355B"
  ) +
  geom_text(
    data = firm_flow_data %>% filter(exit > 0),
    aes(x = x_pos + 0.22, y = -exit, label = comma(exit)),
    vjust = 1.5, size = 3.3, fontface = "bold", colour = "#E76F51"
  ) +
  scale_x_continuous(breaks = firm_flow_data$x_pos, labels = firm_flow_data$survey_year) +
  scale_y_continuous(
    name   = "Number of firms",
    labels = comma_format(),
    expand = expansion(mult = c(0.15, 0.15)),
    sec.axis = sec_axis(
      ~ (. - firm_axis_min) / scale_factor_v3 + index_min_plot,
      name   = "Output index (2019 = 100)",
      labels = number_format(accuracy = 1)
    )
  ) +
  scale_fill_manual(
    values = c("Staying (incumbents)" = "#1B9E77", "Entry" = "#66A61E", "Exit" = "#E76F51"),
    name = "Firm flow"
  ) +
  scale_colour_manual(
    values = c("Output index (2019 = 100)" = "#F4A261"),
    name = "Trend line"
  ) +
  labs(
    title    = glue("{target_sector}: Firm Entry and Exit Dynamics"),
    subtitle = paste(
      "Bars show incumbents, entrants and exits; numbers on bars show total active firms and exit counts.",
      "\nLeft axis measures number of firms. Right axis measures output index (2019 = 100)."
    ),
    x       = NULL,
    caption = paste(
      "Staying and entry bars are stacked above zero. Exit bars are shown separately below zero.",
      "\nThe orange line (right axis) tracks sector real revenue relative to 2019, where 2019 = 100."
    )
  ) +
  guides(
    fill   = guide_legend(order = 1),
    colour = guide_legend(order = 2, override.aes = list(linewidth = 1.25, size = 3))
  ) +
  theme_sector() +
  theme(
    legend.position    = "bottom",
    legend.box         = "vertical",
    legend.title       = element_text(face = "bold"),
    axis.title.y.left  = element_text(face = "bold", colour = "#12355B"),
    axis.text.y.left   = element_text(colour = "#12355B"),
    axis.title.y.right = element_text(face = "bold", colour = "#F4A261"),
    axis.text.y.right  = element_text(colour = "#F4A261")
  )
```

```{r page1-visual3-interpretation, results='asis'}
firm_2023 <- firm_flow_data %>% filter(survey_year == snapshot_year)

entry_exit_msg <- if_else(
  firm_2023$entry > firm_2023$exit,
  "Entry exceeded exit, suggesting positive firm renewal and sector attractiveness.",
  "Exit exceeded entry, indicating restructuring pressure or consolidation within the sector."
)

output_direction <- if_else(
  firm_2023$output_index >= 100,
  glue("expanded to an index of **{round(firm_2023$output_index, 1)}** (2019 = 100)"),
  glue("contracted to an index of **{round(firm_2023$output_index, 1)}** (2019 = 100)")
)

# Renewal-vs-output narrative: correlate firm count with output index across all years
corr_firm_output <- tryCatch(
  cor(firm_flow_data$total_active_firms, firm_flow_data$output_index, use = "complete.obs"),
  error = function(e) NA_real_
)

avg_annual_churn_v3 <- firm_flow_data %>%
  filter(survey_year > base_year) %>%
  summarise(churn = mean(safe_divide(entry + exit, total_active_firms), na.rm = TRUE)) %>%
  pull(churn)

high_churn_v3 <- !is.na(avg_annual_churn_v3) && avg_annual_churn_v3 >= 0.10
pos_corr_v3   <- !is.na(corr_firm_output) && corr_firm_output > 0.5
neg_corr_v3   <- !is.na(corr_firm_output) && corr_firm_output < -0.5

renewal_narrative <- case_when(
  pos_corr_v3 & high_churn_v3 ~
    "Across the period, firm count and output moved together with high churn, consistent with renewal driving output: new entrants reshaped the productive base while aggregate revenue expanded.",
  pos_corr_v3 & !high_churn_v3 ~
    "Firm count and output moved together against a broadly stable firm population, consistent with incumbent-led growth rather than renewal-driven expansion.",
  neg_corr_v3 ~
    "Firm count and output moved in opposite directions across the period, suggesting restructuring: consolidation among fewer, more productive firms sustained performance while total firm numbers declined.",
  is.na(corr_firm_output) ~
    "The relationship between firm population dynamics and aggregate output could not be assessed with available data.",
  TRUE ~
    "Firm count and output moved largely independently, suggesting the link between firm renewal and aggregate performance was weak across the period."
)

interp(glue(
  "In {snapshot_year}, {target_sector} had **{comma(firm_2023$total_active_firms)}** active firms. ",
  "**{comma(firm_2023$entry)}** firms entered and **{comma(firm_2023$exit)}** exited. {entry_exit_msg} ",
  "Sector output {output_direction}. ",
  "{renewal_narrative}"
))
```

## 1.5 Labour Inclusion and Productivity Performance

<p class="section-desc">

Labour productivity and female labour share benchmarked against all-sector averages, `r base_year`-`r snapshot_year`. Left axis: labour productivity (million VND per worker). Right axis: female labour share.

</p>

```{r page1-visual4-prep, include=FALSE}
v4_all_sector_female <- df_page1 %>%
  group_by(survey_year) %>%
  summarise(
    all_female_labour            = sum(labour_female_headcount_end, na.rm = TRUE),
    all_labour                   = sum(labour_headcount_end, na.rm = TRUE),
    female_labour_share_all_sector = safe_divide(all_female_labour, all_labour),
    .groups = "drop"
  )

v4_all_sector_productivity <- all_sector_year %>%
  group_by(survey_year) %>%
  summarise(average_sector_productivity = mean(labour_productivity, na.rm = TRUE), .groups = "drop")

v4_plot_data <- sector_year %>%
  select(survey_year, female_labour_share, labour_productivity) %>%
  left_join(v4_all_sector_female, by = "survey_year") %>%
  left_join(v4_all_sector_productivity, by = "survey_year")

productivity_sector_label  <- glue("Productivity - {target_sector}")
productivity_average_label <- "Average Productivity - All Sectors"
female_sector_label        <- glue("Female Labour Share - {target_sector}")
female_all_label           <- "Female Labour Share - All Sectors"

v4_bar <- v4_plot_data %>%
  transmute(
    survey_year,
    !!productivity_sector_label  := labour_productivity,
    !!productivity_average_label := average_sector_productivity
  ) %>%
  pivot_longer(
    cols = c(all_of(productivity_sector_label), all_of(productivity_average_label)),
    names_to = "Productivity benchmark", values_to = "Productivity"
  ) %>%
  mutate(
    `Productivity benchmark` = factor(`Productivity benchmark`,
      levels = c(productivity_sector_label, productivity_average_label))
  )

v4_line <- v4_plot_data %>%
  transmute(
    survey_year,
    !!female_sector_label := female_labour_share,
    !!female_all_label    := female_labour_share_all_sector
  ) %>%
  pivot_longer(
    cols = c(all_of(female_sector_label), all_of(female_all_label)),
    names_to = "Female labour benchmark", values_to = "Female labour share"
  ) %>%
  mutate(
    `Female labour benchmark` = factor(`Female labour benchmark`,
      levels = c(female_sector_label, female_all_label))
  )

prod_min      <- 0
prod_max      <- max(v4_bar$Productivity, na.rm = TRUE) * 1.22
female_min    <- min(v4_line$`Female labour share`, na.rm = TRUE)
female_max    <- max(v4_line$`Female labour share`, na.rm = TRUE)
female_padding <- (female_max - female_min) * 0.20
female_min_plot <- max(0, female_min - female_padding)
female_max_plot <- min(1, female_max + female_padding)
scale_factor_v4 <- (prod_max - prod_min) / (female_max_plot - female_min_plot)

v4_line <- v4_line %>%
  mutate(female_share_scaled = (`Female labour share` - female_min_plot) * scale_factor_v4 + prod_min)
```

```{r page1-visual4, fig.cap=paste0('Figure 1.4: Labour productivity (million VND per worker) and female labour share benchmarked against all-sector averages, ', base_year, '–', snapshot_year, '.')}
ggplot() +
  geom_col(
    data = v4_bar,
    aes(x = factor(survey_year), y = Productivity, fill = `Productivity benchmark`),
    position = position_dodge(width = 0.7), width = 0.58, colour = "white", linewidth = 0.35
  ) +
  geom_line(
    data = v4_line,
    aes(x = factor(survey_year), y = female_share_scaled,
        colour = `Female labour benchmark`, group = `Female labour benchmark`),
    linewidth = 1.25
  ) +
  geom_point(
    data = v4_line,
    aes(x = factor(survey_year), y = female_share_scaled,
        colour = `Female labour benchmark`, shape = `Female labour benchmark`),
    size = 3
  ) +
  geom_text(
    data = v4_bar,
    aes(x = factor(survey_year), y = Productivity,
        label = comma(round(Productivity, 0)),
        group = `Productivity benchmark`),
    position = position_dodge(width = 0.7),
    vjust = -0.45, size = 3, fontface = "bold", colour = "#12355B"
  ) +
  scale_y_continuous(
    name   = "Labour productivity (million VND per worker)",
    labels = comma_format(accuracy = 1),
    limits = c(prod_min, prod_max),
    sec.axis = sec_axis(
      ~ (. - prod_min) / scale_factor_v4 + female_min_plot,
      name   = "Female labour share",
      labels = percent_format(accuracy = 1)
    )
  ) +
  scale_fill_manual(
    values = setNames(c("#2A9D8F", "#D9E1E8"), c(productivity_sector_label, productivity_average_label)),
    name = "Productivity benchmark", drop = FALSE
  ) +
  scale_colour_manual(
    values = setNames(c("#E76F51", "#6B7280"), c(female_sector_label, female_all_label)),
    name = "Female labour benchmark", drop = FALSE
  ) +
  scale_shape_manual(
    values = setNames(c(16, 15), c(female_sector_label, female_all_label)),
    name = "Female labour benchmark"
  ) +
  labs(
    title    = glue("{target_sector}: Labour Inclusion and Productivity Performance"),
    subtitle = paste(
      glue("Bars compare {target_sector} productivity with average all-sector productivity; lines compare female labour share"),
      "\nLeft axis measures labour productivity. Right axis measures female labour share."
    ),
    x = NULL
  ) +
  guides(
    fill   = guide_legend(order = 1),
    colour = guide_legend(order = 2, override.aes = list(linewidth = 1.3, size = 3)),
    shape  = guide_legend(order = 2)
  ) +
  theme_sector() +
  theme(
    legend.position    = "bottom",
    legend.box         = "vertical",
    legend.title       = element_text(face = "bold"),
    axis.title.y.left  = element_text(face = "bold", colour = "#2A9D8F"),
    axis.text.y.left   = element_text(colour = "#2A9D8F"),
    axis.title.y.right = element_text(face = "bold", colour = "#E76F51"),
    axis.text.y.right  = element_text(colour = "#E76F51")
  )
```

```{r page1-visual4-interpretation, results='asis'}
v4_2019      <- v4_plot_data %>% filter(survey_year == base_year)
v4_2023      <- v4_plot_data %>% filter(survey_year == snapshot_year)
prod_gap     <- v4_2023$labour_productivity - v4_2023$average_sector_productivity
prod_gap_pct <- safe_divide(prod_gap, v4_2023$average_sector_productivity)
prod_dir     <- if_else(prod_gap >= 0, "above", "below")
female_gap   <- v4_2023$female_labour_share - v4_2023$female_labour_share_all_sector
female_dir   <- if_else(female_gap >= 0, "higher than", "lower than")
female_gap_pp <- abs(female_gap) * 100

# LP gap trend: base year vs snapshot year
prod_gap_base <- if (nrow(v4_2019) > 0) {
  v4_2019$labour_productivity - v4_2019$average_sector_productivity
} else { NA_real_ }

prod_trend_msg <- case_when(
  is.na(prod_gap_base) | is.na(prod_gap) ~ "",
  prod_gap > 0 & prod_gap_base <= 0 ~
    "The sector has moved from below to above the economy average in productivity since the base year, recovering a deficit recorded in the base year.",
  prod_gap > 0 & prod_gap_base > 0 & prod_gap > prod_gap_base ~
    "The sector's productivity advantage over the economy average has strengthened since the base year.",
  prod_gap > 0 & prod_gap <= prod_gap_base ~
    "The sector retains a productivity advantage but it has narrowed relative to the base year.",
  prod_gap < 0 & prod_gap > prod_gap_base ~
    "The sector remains below the economy average but has been closing the gap, indicating catch-up dynamics.",
  prod_gap < 0 & prod_gap <= prod_gap_base ~
    "The sector's productivity shortfall relative to the economy average has deepened since the base year.",
  TRUE ~ ""
)

# 4-branch inclusion quadrant
inclusion_msg <- case_when(
  prod_gap >= 0 & female_gap >= 0 ~
    "Above-average productivity alongside above-average female labour share suggests the sector combines economic performance with broad workforce participation , a favourable combination from both efficiency and inclusion perspectives.",
  prod_gap >= 0 & female_gap < 0 ~
    "Above-average productivity but below-average female labour share may reflect capital- or skill-intensive activity with limited workforce inclusion. Policies that widen access without diluting capability could address both dimensions.",
  prod_gap < 0 & female_gap >= 0 ~
    "Below-average productivity despite above-average female labour inclusion suggests employment concentration in lower-productivity activities; capability upgrading is the more binding lever than further workforce expansion.",
  TRUE ~
    "Below-average performance on both productivity and female inclusion indicates a sector where capability investment and active workforce participation policy are both warranted."
)

interp(glue(
  "In {snapshot_year}, {target_sector} labour productivity was **{comma(round(v4_2023$labour_productivity, 1))} million VND per worker**, ",
  "{prod_dir} the all-sector average by **{percent(abs(prod_gap_pct), accuracy = 0.1)}** ",
  "(**{comma(round(v4_2023$average_sector_productivity, 1))} million VND** all-sector). ",
  "{prod_trend_msg} ",
  "Female labour share stood at **{percent(v4_2023$female_labour_share, accuracy = 0.1)}**, ",
  "{female_dir} the all-sector benchmark (**{percent(v4_2023$female_labour_share_all_sector, accuracy = 0.1)}**) ",
  "by **{round(female_gap_pp, 1)} percentage points**. {inclusion_msg}"
))
```

::: part-divider
<span>Part II: Technology Adoption and Innovation Dynamics</span>
:::

# 2. Technology Adoption and Innovation Dynamics

## 2.1 KPI Snapshot

<p class="section-desc">

Key investment and innovation indicators for `r snapshot_year` vs `r base_year`, covering total investment, innovation expenditure, firm participation rates, and labour productivity.

</p>

```{r page2-kpi, results='asis'}
kpi_base_2 <- sector_year %>%
  filter(survey_year == base_year) %>%
  summarise(
    total_investment_base     = sum(total_investment, na.rm = TRUE),
    innovation_investment_base = sum(innovation_investment, na.rm = TRUE),
    labour_productivity_base  = safe_divide(sum(revenue, na.rm = TRUE), sum(labour, na.rm = TRUE)),
    .groups = "drop"
  )

kpi_snapshot_raw_2 <- sector_year %>%
  filter(survey_year == snapshot_year) %>%
  summarise(
    total_investment        = sum(total_investment, na.rm = TRUE),
    innovation_investment   = sum(innovation_investment, na.rm = TRUE),
    firms                   = sum(firms, na.rm = TRUE),
    firms_with_any_investment = sum(firms_with_any_selected_investment, na.rm = TRUE),
    labour_productivity     = safe_divide(sum(revenue, na.rm = TRUE), sum(labour, na.rm = TRUE)),
    .groups = "drop"
  ) %>%
  bind_cols(kpi_base_2) %>%
  mutate(
    total_investment_growth     = safe_divide(total_investment, total_investment_base) - 1,
    innovation_investment_growth = safe_divide(innovation_investment, innovation_investment_base) - 1,
    labour_productivity_growth  = safe_divide(labour_productivity, labour_productivity_base) - 1,
    any_investment_share        = safe_divide(firms_with_any_investment, firms)
  )

kpi_snapshot_2 <- tibble(
  KPI = c("Total investment", "Firms with any investment", "Labour productivity", "Total innovation investment"),
  Value = c(
    paste0(comma(round(kpi_snapshot_raw_2$total_investment / 1e6, 2)), " trillion VND (", format_change(kpi_snapshot_raw_2$total_investment_growth), " vs 2019)"),
    paste0(percent(kpi_snapshot_raw_2$any_investment_share, accuracy = 0.1), " of firms (", comma(kpi_snapshot_raw_2$firms_with_any_investment), " of ", comma(kpi_snapshot_raw_2$firms), " firms)"),
    paste0(comma(round(kpi_snapshot_raw_2$labour_productivity, 1)), " mn VND/worker (", format_change(kpi_snapshot_raw_2$labour_productivity_growth), " vs 2019)"),
    paste0(comma(round(kpi_snapshot_raw_2$innovation_investment / 1e6, 2)), " trillion VND (", format_change(kpi_snapshot_raw_2$innovation_investment_growth), " vs 2019)")
  )
)

kpi_snapshot_2 %>%
  kable(
    format = "html", escape = FALSE,
    caption = glue("{target_sector}: Page 2 KPI Snapshot, {snapshot_year}")
  ) %>%
  kable_styling(
    full_width = TRUE,
    bootstrap_options = c("striped", "hover", "condensed"),
    font_size = 14
  ) %>%
  column_spec(1, bold = TRUE, color = "#12355B", width = "35%") %>%
  column_spec(2, color = "#1F2933", width = "65%")
```

```{r page2-kpi-interpretation, results='asis'}
innov_share_of_total <- safe_divide(kpi_snapshot_raw_2$innovation_investment, kpi_snapshot_raw_2$total_investment)

# Benchmark innovation share against all-sector average in snapshot year
all_innov_share_snap <- df_final %>%
  filter(survey_year == snapshot_year) %>%
  summarise(
    all_innov = sum(innovation_investment_proxy, na.rm = TRUE),
    all_total = sum(total_investment, na.rm = TRUE)
  ) %>%
  summarise(share = safe_divide(all_innov, all_total)) %>%
  pull(share)

innov_benchmark_msg <- case_when(
  is.na(innov_share_of_total) | is.na(all_innov_share_snap) ~
    "The all-sector innovation share benchmark is unavailable.",
  innov_share_of_total > all_innov_share_snap + 0.02 ~
    glue("This is above the all-sector average of **{percent(all_innov_share_snap, accuracy = 0.1)}**, indicating above-average emphasis on capability-led investment relative to the broader economy."),
  innov_share_of_total < all_innov_share_snap - 0.02 ~
    glue("This is below the all-sector average of **{percent(all_innov_share_snap, accuracy = 0.1)}**, suggesting the sector remains more reliant on physical capital than the broader economy."),
  TRUE ~
    glue("This is broadly in line with the all-sector average of **{percent(all_innov_share_snap, accuracy = 0.1)}**.")
)

interp(glue(
  "In {snapshot_year}, total investment in {target_sector} reached ",
  "**{comma(round(kpi_snapshot_raw_2$total_investment / 1e6, 2))} trillion VND** ",
  "({format_change(kpi_snapshot_raw_2$total_investment_growth)} vs {base_year}). ",
  "**{percent(kpi_snapshot_raw_2$any_investment_share, accuracy = 0.1)}** of firms ",
  "(**{comma(kpi_snapshot_raw_2$firms_with_any_investment)}** of **{comma(kpi_snapshot_raw_2$firms)}**) ",
  "undertook at least one form of selected investment. ",
  "Innovation investment reached **{comma(round(kpi_snapshot_raw_2$innovation_investment / 1e6, 2))} trillion VND**, ",
  "representing **{percent(innov_share_of_total, accuracy = 0.1)}** of total investment. ",
  "{innov_benchmark_msg}"
))
```

## 2.2 Investment Scale and Adoption Diffusion

<p class="section-desc">

Total investment volume and the share of firms investing in equipment and innovation, `r base_year`–`r snapshot_year`. Left axis: investment value (trillion VND). Right axis: share of firms investing.

</p>

```{r page2-visual1-prep, include=FALSE}
investment_levels <- c("Equipment investment", "Innovation Investment")

v1_dual <- sector_year %>%
  select(survey_year,
    `Equipment investment`  = equipment_investment,
    `Innovation Investment` = innovation_investment
  ) %>%
  pivot_longer(cols = -survey_year, names_to = "Investment type", values_to = "Investment value") %>%
  mutate(
    `Investment type`   = factor(`Investment type`, levels = investment_levels),
    investment_trillion = `Investment value` / 1e6
  )

v1_rate_dual <- sector_year %>%
  select(survey_year,
    `Equipment investment`  = equipment_adoption_rate,
    `Innovation Investment` = innovation_adoption_rate
  ) %>%
  pivot_longer(cols = -survey_year, names_to = "Investment type", values_to = "Adoption rate") %>%
  mutate(`Investment type` = factor(`Investment type`, levels = investment_levels))

max_investment <- max(v1_dual$investment_trillion, na.rm = TRUE)
max_rate       <- max(v1_rate_dual$`Adoption rate`, na.rm = TRUE)
scale_factor   <- ifelse(max_rate > 0, max_investment / max_rate, 1)
```

```{r page2-visual1, fig.cap=paste0('Figure 2.1: Total equipment and innovation investment value (trillion VND) and share of firms investing, ', base_year, '–', snapshot_year, '.')}
ggplot() +
  geom_col(
    data = v1_dual,
    aes(x = factor(survey_year), y = investment_trillion, fill = `Investment type`),
    position = position_dodge(width = 0.72), width = 0.62, colour = "white", linewidth = 0.3,
    show.legend = TRUE
  ) +
  geom_line(
    data = v1_rate_dual,
    aes(x = factor(survey_year), y = `Adoption rate` * scale_factor,
        colour = `Investment type`, group = `Investment type`),
    linewidth = 0.85, show.legend = TRUE
  ) +
  geom_point(
    data = v1_rate_dual,
    aes(x = factor(survey_year), y = `Adoption rate` * scale_factor, colour = `Investment type`),
    size = 2.5, show.legend = TRUE
  ) +
  geom_text(
    data = v1_dual,
    aes(x = factor(survey_year), y = investment_trillion,
        label = comma(round(investment_trillion, 2)),
        group = `Investment type`),
    position = position_dodge(width = 0.72),
    vjust = -0.45, size = 3, fontface = "bold", colour = "#12355B"
  ) +
  scale_y_continuous(
    name   = "Total investment value (trillion VND)",
    labels = comma_format(accuracy = 0.01),
    expand = expansion(mult = c(0, 0.15)),
    sec.axis = sec_axis(
      ~ . / scale_factor,
      name   = "Share of firms investing",
      labels = percent_format(accuracy = 1)
    )
  ) +
  scale_fill_manual(
    values = c("Equipment investment" = "#2A9D8F", "Innovation Investment" = "#F4A261"),
    name = "Investment value", drop = FALSE
  ) +
  scale_colour_manual(
    values = c("Equipment investment" = "#12355B", "Innovation Investment" = "#E76F51"),
    name = "Adoption rate", drop = FALSE
  ) +
  labs(
    title    = glue("{target_sector}: Investment Scale and Adoption Diffusion"),
    subtitle = paste(
      "Bars show total investment value; lines show share of firms investing",
      "\nLeft axis measures total investment value (trillion VND). Right axis measures the share of firms investing."
    ),
    x = NULL
  ) +
  guides(
    fill   = guide_legend(title = "Investment value", order = 1),
    colour = guide_legend(title = "Adoption rate", order = 2, override.aes = list(linewidth = 1.1, size = 2.8))
  ) +
  theme_sector() +
  theme(
    legend.position    = "bottom",
    legend.box         = "vertical",
    legend.title       = element_text(face = "bold"),
    axis.title.y.left  = element_text(face = "bold", colour = "#12355B"),
    axis.text.y.left   = element_text(colour = "#12355B"),
    axis.title.y.right = element_text(face = "bold", colour = "#E76F51"),
    axis.text.y.right  = element_text(colour = "#E76F51")
  )
```

```{r page2-visual1-interpretation, results='asis'}
equip_rate_2023 <- v1_rate_dual %>%
  filter(survey_year == snapshot_year, `Investment type` == "Equipment investment") %>%
  pull(`Adoption rate`)
innov_rate_2023 <- v1_rate_dual %>%
  filter(survey_year == snapshot_year, `Investment type` == "Innovation Investment") %>%
  pull(`Adoption rate`)
equip_val_2023  <- v1_dual %>%
  filter(survey_year == snapshot_year, `Investment type` == "Equipment investment") %>%
  pull(investment_trillion)
innov_val_2023  <- v1_dual %>%
  filter(survey_year == snapshot_year, `Investment type` == "Innovation Investment") %>%
  pull(investment_trillion)

adoption_gap_pp_v1 <- abs(equip_rate_2023 - innov_rate_2023) * 100

adoption_gap_msg <- case_when(
  adoption_gap_pp_v1 <= 5 ~
    "Equipment and innovation adoption rates are closely matched, suggesting firms treat the two investment types as complementary rather than substitutes.",
  equip_rate_2023 > innov_rate_2023 + 0.05 ~
    glue("Equipment adoption is materially higher than innovation adoption ({percent(equip_rate_2023, accuracy = 0.1)} vs {percent(innov_rate_2023, accuracy = 0.1)}), consistent with a sector still in an earlier upgrading phase where hardware investment leads capability development."),
  innov_rate_2023 > equip_rate_2023 + 0.05 ~
    glue("Innovation adoption is materially higher than equipment adoption ({percent(innov_rate_2023, accuracy = 0.1)} vs {percent(equip_rate_2023, accuracy = 0.1)}), suggesting firms are emphasising intangible capability ahead of physical hardware , an advanced adoption profile."),
  TRUE ~ ""
)

diffusion_level_msg <- case_when(
  max(equip_rate_2023, innov_rate_2023) < 0.10 ~
    "Overall adoption remains very limited, with the large majority of firms not investing in either category.",
  max(equip_rate_2023, innov_rate_2023) < 0.20 ~
    "Adoption is still in early diffusion across both categories, with most firms not yet participating in either type of investment.",
  max(equip_rate_2023, innov_rate_2023) < 0.40 ~
    "Adoption has reached moderate breadth in at least one category, suggesting technology investment is beginning to move beyond early adopters.",
  TRUE ~
    "Adoption is relatively broad-based across at least one investment type, indicating technology investment has diffused well beyond a narrow group of early adopters."
)

interp(glue(
  "In {snapshot_year}, equipment investment was adopted by **{percent(equip_rate_2023, accuracy = 0.1)}** of firms ",
  "(**{comma(round(equip_val_2023, 2))} trillion VND** total), while innovation investment was adopted by ",
  "**{percent(innov_rate_2023, accuracy = 0.1)}** of firms (**{comma(round(innov_val_2023, 2))} trillion VND** total). ",
  "{adoption_gap_msg} ",
  "{diffusion_level_msg}"
))
```

## 2.3 Investment Allocation Mix and Total Investment

<p class="section-desc">

Stacked shares of investment categories (machinery, innovation, construction, repairs) and total investment volume, `r base_year`–`r snapshot_year`. Left axis: allocation share. Right axis: total investment (trillion VND).

</p>

```{r page2-visual2-prep, include=FALSE}
v2_plot_data <- sector_df %>%
  filter(has_equipment_investment == 1 | has_innovation_investment == 1 | total_investment > 0) %>%
  group_by(survey_year) %>%
  summarise(
    `Machinery and equipment`      = sum(investment_machinery_equipment, na.rm = TRUE),
    `Innovation Investment`        = sum(innovation_investment_proxy, na.rm = TRUE),
    `Construction and installation` = sum(investment_construction_installation, na.rm = TRUE),
    `Other construction`           = sum(investment_construction_other, na.rm = TRUE),
    `Major repairs`                = sum(investment_major_repairs, na.rm = TRUE),
    .groups = "drop"
  )

v2_mix <- v2_plot_data %>%
  pivot_longer(-survey_year, names_to = "Investment category", values_to = "Value") %>%
  group_by(survey_year) %>%
  mutate(
    total_value = sum(Value, na.rm = TRUE),
    share       = if_else(total_value > 0, Value / total_value, NA_real_)
  ) %>%
  ungroup() %>%
  filter(!is.na(share))

v2_total <- v2_plot_data %>%
  mutate(
    total_investment_selected = `Machinery and equipment` + `Innovation Investment` +
      `Construction and installation` + `Other construction` + `Major repairs`,
    total_investment_trillion = total_investment_selected / 1e6
  ) %>%
  select(survey_year, total_investment_trillion)

scale_factor_v2 <- max(v2_total$total_investment_trillion, na.rm = TRUE)
```

```{r page2-visual2, fig.cap=paste0('Figure 2.2: Investment allocation shares by category and total selected investment volume (trillion VND), ', base_year, '–', snapshot_year, '.')}
ggplot() +
  geom_col(
    data = v2_mix,
    aes(x = factor(survey_year), y = share, fill = `Investment category`),
    width = 0.72, colour = "white", linewidth = 0.28
  ) +
  geom_line(
    data = v2_total,
    aes(x = factor(survey_year), y = total_investment_trillion / scale_factor_v2,
        group = 1, colour = "Total investment"),
    linewidth = 1.3
  ) +
  geom_point(
    data = v2_total,
    aes(x = factor(survey_year), y = total_investment_trillion / scale_factor_v2,
        colour = "Total investment"),
    size = 3
  ) +
  scale_y_continuous(
    name   = "Investment allocation share",
    labels = percent_format(accuracy = 1),
    sec.axis = sec_axis(
      ~ . * scale_factor_v2,
      name   = "Total selected investment (trillion VND)",
      labels = comma_format(accuracy = 0.01)
    )
  ) +
  scale_fill_manual(
    values = c(
      "Machinery and equipment"       = "#2A9D8F",
      "Innovation Investment"         = "#F4A261",
      "Construction and installation" = "#8ECAE6",
      "Other construction"            = "#B8C0CC",
      "Major repairs"                 = "#E9C46A"
    ),
    name = "Investment category", drop = FALSE
  ) +
  scale_colour_manual(values = c("Total investment" = "#12355B"), name = "Total investment line") +
  labs(
    title    = glue("{target_sector}: Investment Allocation Mix and Total Investment"),
    subtitle = paste(
      "Bars show allocation shares; line shows total selected investment value",
      "\nLeft axis measures investment allocation share. Right axis measures total selected investment value (trillion VND)."
    ),
    x = NULL
  ) +
  guides(
    fill   = guide_legend(order = 1, override.aes = list(alpha = 1)),
    colour = guide_legend(order = 2, override.aes = list(linewidth = 1.3, size = 3))
  ) +
  theme_sector() +
  theme(
    legend.position    = "bottom",
    legend.box         = "vertical",
    legend.title       = element_text(face = "bold"),
    plot.subtitle      = element_text(size = 10.5, colour = "#6B7280", lineheight = 1.25),
    axis.title.y.left  = element_text(face = "bold", colour = "#2A9D8F"),
    axis.text.y.left   = element_text(colour = "#2A9D8F"),
    axis.title.y.right = element_text(face = "bold", colour = "#12355B"),
    axis.text.y.right  = element_text(colour = "#12355B")
  )
```

```{r page2-visual2-interpretation, results='asis'}
dominant_investment  <- v2_mix %>% filter(survey_year == snapshot_year) %>% arrange(desc(share)) %>% slice(1)
total_selected_2023  <- v2_total %>% filter(survey_year == snapshot_year) %>% pull(total_investment_trillion)

innov_share_base     <- v2_mix %>%
  filter(survey_year == base_year, `Investment category` == "Innovation Investment") %>%
  pull(share)
innov_share_snap     <- v2_mix %>%
  filter(survey_year == snapshot_year, `Investment category` == "Innovation Investment") %>%
  pull(share)
innov_share_base     <- if (length(innov_share_base) > 0) innov_share_base else NA_real_
innov_share_snap     <- if (length(innov_share_snap) > 0) innov_share_snap else NA_real_
innov_trend_pp       <- (innov_share_snap - innov_share_base) * 100
innov_trend_dir      <- if_else(!is.na(innov_trend_pp) && innov_trend_pp >= 0, "rose", "fell")

investment_mix_msg <- case_when(
  dominant_investment$`Investment category` == "Machinery and equipment" ~
    "This indicates a stronger focus on hard-technology adoption and physical capital upgrading.",
  dominant_investment$`Investment category` == "Innovation Investment" ~
    "This indicates a stronger emphasis on innovation capability, R&D, and soft-technology investment.",
  TRUE ~
    "Investment is directed primarily toward infrastructure, construction, or maintenance activities."
)

innov_trend_sentence <- if_else(
  !is.na(innov_trend_pp),
  glue(
    "The innovation investment share {innov_trend_dir} from **{percent(innov_share_base, accuracy = 0.1)}** ",
    "in {base_year} to **{percent(innov_share_snap, accuracy = 0.1)}** in {snapshot_year} ",
    "({if_else(innov_trend_pp >= 0, '+', '')}**{round(innov_trend_pp, 1)} percentage points**), ",
    "indicating {if_else(innov_trend_pp >= 0, 'a growing', 'a shrinking')} role for innovation ",
    "relative to physical capital."
  ),
  ""
)

interp(glue(
  "In {snapshot_year}, total selected investment reached **{comma(round(total_selected_2023, 2))} trillion VND**. ",
  "The largest allocation was {dominant_investment$`Investment category`} ",
  "at **{percent(dominant_investment$share, accuracy = 0.1)}**. {investment_mix_msg} ",
  "{innov_trend_sentence}"
))
```

## 2.4 Frontier vs Laggard Gap over Time

<p class="section-desc">

Productivity and investment intensity ratios between the top 3% of firms (frontier) and the remaining 97% (laggards), showing whether the performance gap is widening or narrowing.

</p>

```{r page2-visual3-prep, include=FALSE}
frontier_year_base <- sector_df %>%
  mutate(firm_productivity = safe_divide(total_net_revenue, labour_headcount_end)) %>%
  filter(labour_headcount_end > 0, firm_productivity > 0) %>%
  group_by(survey_year) %>%
  mutate(
    productivity_rank = percent_rank(firm_productivity),
    frontier_group    = if_else(productivity_rank >= 0.97, "Frontier: top 3%", "Laggard: remaining 97%")
  ) %>%
  ungroup()

frontier_year_metrics <- frontier_year_base %>%
  group_by(survey_year, frontier_group) %>%
  summarise(
    `Labour productivity`              = median(firm_productivity, na.rm = TRUE),
    `Equipment investment per worker`  = safe_divide(sum(investment_machinery_equipment, na.rm = TRUE), sum(labour_headcount_end, na.rm = TRUE)),
    `Innovation investment per worker` = safe_divide(sum(innovation_investment_proxy, na.rm = TRUE), sum(labour_headcount_end, na.rm = TRUE)),
    .groups = "drop"
  ) %>%
  pivot_longer(
    cols = c(`Labour productivity`, `Equipment investment per worker`, `Innovation investment per worker`),
    names_to = "Metric", values_to = "Value"
  ) %>%
  pivot_wider(names_from = frontier_group, values_from = Value) %>%
  mutate(
    Ratio = safe_divide(`Frontier: top 3%`, `Laggard: remaining 97%`),
    Label = paste0(round(Ratio, 1), "x")
  ) %>%
  filter(is.finite(Ratio)) %>%
  mutate(
    Metric = factor(Metric, levels = c(
      "Labour productivity",
      "Equipment investment per worker",
      "Innovation investment per worker"
    ))
  )
```

```{r page2-visual3, fig.cap=paste0('Figure 2.3: Frontier-to-laggard ratio for labour productivity and investment intensity (top 3% vs remaining 97% of firms), ', base_year, '–', snapshot_year, '.')}
ggplot(frontier_year_metrics, aes(x = survey_year, y = Ratio, colour = Metric, group = Metric)) +
  geom_hline(yintercept = 1, colour = "#B8C0CC", linewidth = 0.8) +
  geom_line(linewidth = 1.4) +
  geom_point(size = 3.8, stroke = 0.9, shape = 21, fill = "white") +
  geom_text(
    aes(label = Label), vjust = -0.9, size = 3.4, fontface = "bold",
    colour = "#12355B", show.legend = FALSE
  ) +
  scale_x_continuous(breaks = sort(unique(frontier_year_metrics$survey_year))) +
  scale_y_continuous(
    labels = number_format(accuracy = 0.1),
    expand = expansion(mult = c(0.05, 0.25))
  ) +
  scale_colour_manual(
    values = c(
      "Labour productivity"              = "#E63946",
      "Equipment investment per worker"  = "#2A9D8F",
      "Innovation investment per worker" = "#F4A261"
    ),
    drop = FALSE
  ) +
  guides(colour = guide_legend(title = "Metric", override.aes = list(size = 4, linewidth = 1.5))) +
  labs(
    title    = glue("{target_sector}: Frontier vs Laggard Gap over Time"),
    subtitle = "Top 3% firms vs remaining 97%: multiplier of productivity and investment intensity",
    x       = NULL,
    y       = "Frontier / laggard multiplier",
    colour  = "Metric",
    caption = "Ratio > 1 means frontier firms have higher values than laggard firms."
  ) +
  theme_sector() +
  theme(
    legend.position  = "bottom",
    legend.direction = "horizontal",
    legend.title     = element_text(face = "bold"),
    legend.text      = element_text(size = 10)
  )
```

```{r page2-visual3-interpretation, results='asis'}
frontier_2023 <- frontier_year_metrics %>% filter(survey_year == snapshot_year)
frontier_base <- frontier_year_metrics %>% filter(survey_year == base_year)

prod_ratio      <- frontier_2023 %>% filter(Metric == "Labour productivity") %>% pull(Ratio)
equip_ratio     <- frontier_2023 %>% filter(Metric == "Equipment investment per worker") %>% pull(Ratio)
innov_ratio     <- frontier_2023 %>% filter(Metric == "Innovation investment per worker") %>% pull(Ratio)
prod_ratio_base <- frontier_base %>% filter(Metric == "Labour productivity") %>% pull(Ratio)

prod_ratio_base_val <- if (length(prod_ratio_base) > 0 && !is.na(prod_ratio_base[1])) prod_ratio_base[1] else NA_real_
prod_ratio_val      <- if (length(prod_ratio) > 0 && !is.na(prod_ratio[1])) prod_ratio[1] else NA_real_

gap_trend <- if_else(
  !is.na(prod_ratio_base_val) && !is.na(prod_ratio_val),
  if_else(prod_ratio_val > prod_ratio_base_val, "widened", "narrowed"),
  "could not be trended with available data"
)

trend_detail <- if_else(
  !is.na(prod_ratio_base_val) && !is.na(prod_ratio_val),
  glue("The productivity gap has {gap_trend} since {base_year} (**{round(prod_ratio_base_val, 1)}x** then vs **{round(prod_ratio_val, 1)}x** now). "),
  ""
)

gap_msg <- if_else(
  !is.na(prod_ratio_val) && prod_ratio_val > 10,
  "The large gap points to substantial unevenness, with a small frontier cohort capturing a disproportionate share of sector productivity.",
  "The productivity gap is present but less extreme, suggesting some degree of capability diffusion across the sector."
)

equip_ratio_val <- if (length(equip_ratio) > 0 && !is.na(equip_ratio[1])) equip_ratio[1] else NA_real_
innov_ratio_val <- if (length(innov_ratio) > 0 && !is.na(innov_ratio[1])) innov_ratio[1] else NA_real_

binding_gap_msg <- case_when(
  is.na(equip_ratio_val) | is.na(innov_ratio_val) ~
    "The two diffusion gaps could not be compared with available data.",
  equip_ratio_val > innov_ratio_val + 0.5 ~
    "The larger gap is in equipment intensity, pointing to capital and financing constraints as the more binding barrier when laggard firms attempt to upgrade machinery.",
  innov_ratio_val > equip_ratio_val + 0.5 ~
    "The larger gap is in innovation intensity, pointing to skills, organisational capability and digital readiness as the more binding constraint on laggard firm catch-up.",
  TRUE ~
    "The equipment-intensity and innovation-intensity gaps are of similar magnitude, indicating that capital and capability constraints likely operate together rather than one dominating the other."
)

interp(glue(
  "In {snapshot_year}, frontier firms (top 3%) in {target_sector} were **{round(prod_ratio_val, 1)}x** more productive ",
  "than laggard firms (remaining 97%). Their equipment and innovation investment intensity were ",
  "**{round(equip_ratio_val, 1)}x** and **{round(innov_ratio_val, 1)}x** higher, respectively. ",
  "{trend_detail}{gap_msg} ",
  "{binding_gap_msg}"
))
```

## 2.5 Equipment Investment Cohort Panel

<p class="section-desc">

Equipment investment persistence among balanced-panel firms (observed in all `r length(unique(sector_df$survey_year))` survey years), grouped by the year of first equipment investment.

</p>

```{r page2-visual4-prep, include=FALSE}
# Balanced panel definition: firms observed in every survey year for this sector.
# Restricting to balanced-panel firms ensures that investment persistence is
# measured on the same set of firms across all years, avoiding compositional
# changes caused by entry or exit. Firms that appear in fewer than all years
# are excluded because their absence in some years cannot be distinguished from
# zero investment versus non-response or exit.
all_years <- sort(unique(sector_df$survey_year))
n_years   <- length(all_years)

balanced_firms <- sector_df %>%
  distinct(tax_id, survey_year) %>%
  count(tax_id, name = "n_years_observed") %>%
  filter(n_years_observed == n_years) %>%
  pull(tax_id)

sector_df_balanced <- sector_df %>% filter(tax_id %in% balanced_firms)

firm_years_balanced <- sector_df_balanced %>%
  select(tax_id, survey_year, has_equipment_investment) %>%
  distinct()

first_invest_year_balanced <- firm_years_balanced %>%
  filter(has_equipment_investment == 1) %>%
  group_by(tax_id) %>%
  summarise(first_equipment_year = min(survey_year), .groups = "drop")

min_year <- min(firm_years_balanced$survey_year, na.rm = TRUE)

cohort_df_balanced <- firm_years_balanced %>%
  left_join(first_invest_year_balanced, by = "tax_id") %>%
  mutate(cohort = case_when(
    is.na(first_equipment_year)        ~ "Never investor",
    first_equipment_year <= min_year   ~ paste0("Already investor in ", min_year),
    TRUE                               ~ paste0("First invested in ", first_equipment_year)
  ))

cohort_order_balanced <- cohort_df_balanced %>%
  distinct(cohort, first_equipment_year) %>%
  mutate(order_value = case_when(
    cohort == "Never investor"          ~ 9999,
    str_detect(cohort, "Already investor") ~ 0,
    TRUE                                ~ first_equipment_year
  )) %>%
  arrange(desc(order_value)) %>%
  pull(cohort)

cohort_plot_data_balanced <- cohort_df_balanced %>%
  group_by(cohort, survey_year) %>%
  summarise(
    n_observed  = n_distinct(tax_id),
    n_investing = n_distinct(tax_id[has_equipment_investment == 1]),
    share_investing = safe_divide(n_investing, n_observed),
    .groups = "drop"
  ) %>%
  mutate(
    cohort = factor(cohort, levels = unique(cohort_order_balanced)),
    label  = paste0(percent(share_investing, accuracy = 0.1), "\n", comma(n_investing), " of ", comma(n_observed))
  )
```

```{r page2-visual4, fig.cap=paste0('Figure 2.4: Equipment investment persistence by cohort among balanced-panel firms observed in all ', n_years, ' survey years.')}
ggplot(cohort_plot_data_balanced, aes(x = factor(survey_year), y = cohort, fill = share_investing)) +
  geom_tile(colour = "white", linewidth = 0.6) +
  geom_text(aes(label = label), size = 3.15, fontface = "bold", colour = "#1F2933", lineheight = 0.9) +
  scale_fill_gradient(
    low = "#FFFBF0", high = "#4DB6AC",
    labels = percent_format(accuracy = 1), limits = c(0, 1)
  ) +
  labs(
    title   = glue("{target_sector}: Equipment Investment Cohort Panel"),
    subtitle = glue("Balanced panel only: firms observed in all {n_years} years"),
    x       = "Survey year",
    y       = "Equipment investment cohort",
    fill    = "Share investing",
    caption = glue(
      "Balanced-panel firms only. N = {comma(length(balanced_firms))}. ",
      "Cell labels show percentage and number of investing firms among firms in each cohort-year."
    )
  ) +
  theme_sector() +
  theme(panel.grid = element_blank())
```

```{r page2-visual4-interpretation, results='asis'}
n_balanced    <- length(balanced_firms)
n_total_firms <- n_distinct(sector_df$tax_id)
balanced_share <- safe_divide(n_balanced, n_total_firms)

never_count <- cohort_plot_data_balanced %>%
  filter(str_detect(as.character(cohort), "Never investor"), survey_year == base_year) %>%
  pull(n_observed)
never_share <- safe_divide(
  if (length(never_count) > 0) sum(never_count) else 0,
  n_balanced
)

early_rate_raw <- cohort_plot_data_balanced %>%
  filter(str_detect(as.character(cohort), "Already investor"), survey_year == snapshot_year) %>%
  pull(share_investing)
early_rate <- if (length(early_rate_raw) > 0 && !is.na(early_rate_raw[1])) early_rate_raw[1] else NA_real_

early_msg <- if_else(
  !is.na(early_rate),
  glue(
    "Among firms already investing in {base_year}, **{percent(early_rate, accuracy = 0.1)}** continued ",
    "investing in {snapshot_year}, indicating {if_else(early_rate > 0.5, 'strong', 'moderate')} ",
    "investment persistence."
  ),
  glue(
    "No 'already investor in {base_year}' cohort was observed, suggesting most equipment adoption ",
    "in this sector began after {base_year}."
  )
)

# Representativeness signal
panel_rep_msg <- case_when(
  balanced_share >= 0.60 ~
    "The balanced panel covers the majority of unique firms, so cohort persistence rates are broadly representative of the sector.",
  balanced_share >= 0.35 ~
    "The balanced panel covers a minority of unique firms; results reflect long-tenure incumbents and may understate adoption volatility among transient entrants and exits.",
  TRUE ~
    "Only a small share of all unique firms appear in every survey year; the cohort analysis should be read as descriptive of long-tenure incumbents only."
)

# Latest-cohort adoption trajectory vs early cohort
latest_cohort_rate_raw <- cohort_plot_data_balanced %>%
  filter(str_detect(as.character(cohort), as.character(snapshot_year)), survey_year == snapshot_year) %>%
  pull(share_investing)
latest_cohort_rate <- if (length(latest_cohort_rate_raw) > 0 && !is.na(latest_cohort_rate_raw[1])) latest_cohort_rate_raw[1] else NA_real_

cohort_trajectory_msg <- if (!is.na(latest_cohort_rate) && !is.na(early_rate)) {
  if_else(
    latest_cohort_rate >= early_rate - 0.10,
    "New-entry cohorts maintain adoption rates comparable to earlier cohorts, suggesting the equipment investment market in this sector is not saturating.",
    "New-entry cohorts show lower adoption rates than earlier cohorts, consistent with either higher entry barriers for newer firms or later entrants being structurally less capital-intensive."
  )
} else { "" }

interp(glue(
  "The balanced panel covers **{comma(n_balanced)}** firms observed in all {n_years} survey years ",
  "(**{percent(balanced_share, accuracy = 0.1)}** of all unique firms in this sector). ",
  "{panel_rep_msg} ",
  "**{percent(never_share, accuracy = 0.1)}** of balanced-panel firms never invested in equipment ",
  "across the full period. {early_msg} ",
  "{cohort_trajectory_msg}"
))
```

## 2.6 Provincial Technology Adoption Maps

<p class="section-desc">

Equipment and innovation adoption rates mapped across Vietnam's 34 consolidated provinces for `r base_year` and `r snapshot_year`, showing where technology investment is geographically concentrated.

</p>

```{r page2-visual5-prep, include=FALSE}
target_years <- c(2019, 2023)

province_lookup_old <- tibble(
  province_code    = c(1,2,4,6,8,10,11,12,14,15,17,19,20,22,24,25,26,27,30,31,33,34,35,36,37,38,40,42,44,45,46,48,49,51,52,54,56,58,60,62,64,66,67,68,70,72,74,75,77,79,80,82,83,84,86,87,89,91,92,93,94,95,96),
  old_province_name = c("Ha Noi","Ha Giang","Cao Bang","Bac Kan","Tuyen Quang","Lao Cai","Dien Bien","Lai Chau","Son La","Yen Bai","Hoa Binh","Thai Nguyen","Lang Son","Quang Ninh","Bac Giang","Phu Tho","Vinh Phuc","Bac Ninh","Hai Duong","Hai Phong","Hung Yen","Thai Binh","Ha Nam","Nam Dinh","Ninh Binh","Thanh Hoa","Nghe An","Ha Tinh","Quang Binh","Quang Tri","Thua Thien Hue","Da Nang","Quang Nam","Quang Ngai","Binh Dinh","Phu Yen","Khanh Hoa","Ninh Thuan","Binh Thuan","Kon Tum","Gia Lai","Dak Lak","Dak Nong","Lam Dong","Binh Phuoc","Tay Ninh","Binh Duong","Dong Nai","Ba Ria - Vung Tau","Ho Chi Minh City","Long An","Tien Giang","Ben Tre","Tra Vinh","Vinh Long","Dong Thap","An Giang","Kien Giang","Can Tho","Hau Giang","Soc Trang","Bac Lieu","Ca Mau")
) %>% mutate(old_province_key = clean_name(old_province_name))

province_34_mapping <- province_lookup_old %>%
  mutate(
    new_province_name = case_when(
      old_province_name == "Ha Noi"                                            ~ "Ha Noi",
      old_province_name %in% c("Tuyen Quang", "Ha Giang")                     ~ "Tuyen Quang",
      old_province_name %in% c("Lao Cai", "Yen Bai")                          ~ "Lao Cai",
      old_province_name %in% c("Thai Nguyen", "Bac Kan")                      ~ "Thai Nguyen",
      old_province_name %in% c("Phu Tho", "Vinh Phuc", "Hoa Binh")           ~ "Phu Tho",
      old_province_name %in% c("Bac Ninh", "Bac Giang")                       ~ "Bac Ninh",
      old_province_name %in% c("Hung Yen", "Thai Binh")                       ~ "Hung Yen",
      old_province_name %in% c("Hai Phong", "Hai Duong")                      ~ "Hai Phong",
      old_province_name %in% c("Ninh Binh", "Nam Dinh", "Ha Nam")             ~ "Ninh Binh",
      old_province_name == "Thua Thien Hue"                                   ~ "Hue",
      old_province_name %in% c("Quang Binh", "Quang Tri")                     ~ "Quang Tri",
      old_province_name %in% c("Da Nang", "Quang Nam")                        ~ "Da Nang",
      old_province_name %in% c("Quang Ngai", "Kon Tum")                       ~ "Quang Ngai",
      old_province_name %in% c("Gia Lai", "Binh Dinh")                        ~ "Gia Lai",
      old_province_name %in% c("Khanh Hoa", "Ninh Thuan")                     ~ "Khanh Hoa",
      old_province_name %in% c("Lam Dong", "Binh Thuan", "Dak Nong")          ~ "Lam Dong",
      old_province_name %in% c("Dak Lak", "Phu Yen")                          ~ "Dak Lak",
      old_province_name %in% c("Ho Chi Minh City", "Binh Duong", "Ba Ria - Vung Tau") ~ "Ho Chi Minh City",
      old_province_name %in% c("Dong Nai", "Binh Phuoc")                      ~ "Dong Nai",
      old_province_name %in% c("Tay Ninh", "Long An")                         ~ "Tay Ninh",
      old_province_name %in% c("Can Tho", "Soc Trang", "Hau Giang")           ~ "Can Tho",
      old_province_name %in% c("Vinh Long", "Ben Tre", "Tra Vinh")            ~ "Vinh Long",
      old_province_name %in% c("Dong Thap", "Tien Giang")                     ~ "Dong Thap",
      old_province_name %in% c("Ca Mau", "Bac Lieu")                          ~ "Ca Mau",
      old_province_name %in% c("An Giang", "Kien Giang")                      ~ "An Giang",
      TRUE                                                                     ~ old_province_name
    ),
    new_province_key = clean_name(new_province_name)
  )

province_adoption_34 <- sector_df %>%
  filter(survey_year %in% target_years, !is.na(province_code), !is.na(survey_year)) %>%
  left_join(
    province_34_mapping %>% select(province_code, old_province_name, new_province_name, new_province_key),
    by = "province_code"
  ) %>%
  filter(!is.na(new_province_name)) %>%
  group_by(new_province_name, new_province_key, survey_year) %>%
  summarise(
    total_firms            = n_distinct(tax_id),
    firms_with_equipment   = n_distinct(tax_id[has_equipment_investment == 1]),
    firms_with_innovation  = n_distinct(tax_id[has_innovation_investment == 1]),
    equipment_adoption_rate  = safe_divide(firms_with_equipment, total_firms),
    innovation_adoption_rate = safe_divide(firms_with_innovation, total_firms),
    .groups = "drop"
  )

vn_map_raw <- geodata::gadm(country = "VNM", level = 1, path = tempdir())

vn_map_old <- st_as_sf(vn_map_raw) %>%
  st_make_valid() %>%
  mutate(
    old_province_key = clean_name(NAME_1),
    old_province_key = case_when(
      old_province_key == "hochiminh" ~ "hochiminhcity",
      TRUE ~ old_province_key
    )
  ) %>%
  left_join(
    province_lookup_old %>% select(province_code, old_province_name, old_province_key),
    by = "old_province_key"
  ) %>%
  left_join(
    province_34_mapping %>% select(province_code, new_province_name, new_province_key),
    by = "province_code"
  )

vn_map_34 <- vn_map_old %>%
  filter(!is.na(new_province_name)) %>%
  group_by(new_province_name, new_province_key) %>%
  summarise(
    old_provinces = paste(sort(unique(old_province_name)), collapse = " + "),
    geometry      = st_union(geometry),
    .groups = "drop"
  ) %>%
  st_make_valid()

vn_map_34_expanded <- tidyr::crossing(survey_year = target_years, vn_map_34 %>% st_drop_geometry()) %>%
  left_join(vn_map_34 %>% select(new_province_key, geometry), by = "new_province_key") %>%
  st_as_sf()

vn_adoption_34_map <- vn_map_34_expanded %>%
  left_join(province_adoption_34, by = c("new_province_key", "new_province_name", "survey_year"))

make_island_polygon <- function(lon, lat, radius = 0.06, n = 30) {
  angle  <- seq(0, 2 * pi, length.out = n)
  coords <- cbind(lon + radius * cos(angle), lat + radius * sin(angle))
  st_polygon(list(coords))
}

island_polygons_base <- bind_rows(
  tibble(
    island_group = "Hoang Sa",
    lon    = c(111.2,111.5,111.8,112.1,112.4,112.7,111.7,112.0,112.3,112.9),
    lat    = c(16.1,16.4,16.2,16.5,16.7,16.3,16.8,15.9,16.1,16.6),
    radius = c(0.065,0.055,0.060,0.050,0.055,0.050,0.045,0.045,0.040,0.040)
  ),
  tibble(
    island_group = "Truong Sa",
    lon    = c(110.7,113.1,113.5,113.9,114.3,114.7,115.1,115.5,115.9,116.2,112.9,113.4,113.8,114.4,115.0,115.7),
    lat    = c(10.4,10.8,10.2,10.6,10.0,9.6,10.4,9.8,9.2,9.6,9.3,8.9,8.5,8.8,8.4,8.7),
    radius = c(0.060,0.045,0.060,0.050,0.055,0.050,0.045,0.045,0.040,0.040,0.045,0.040,0.040,0.040,0.035,0.035)
  )
) %>%
  mutate(
    island_id = row_number(),
    geometry  = pmap(list(lon, lat, radius), ~ make_island_polygon(..1, ..2, ..3))
  ) %>%
  st_as_sf(crs = 4326)

island_polygons <- island_polygons_base %>%
  tidyr::crossing(survey_year = target_years) %>%
  st_as_sf()

island_labels <- tibble(
  island_group = c("Hoang Sa", "Truong Sa"),
  lon = c(113.5, 115.0),
  lat = c(16.5, 9.0)
) %>%
  tidyr::crossing(survey_year = target_years) %>%
  st_as_sf(coords = c("lon", "lat"), crs = 4326)
```

```{r page2-visual5, fig.width=12, fig.height=7, fig.cap=paste0('Figure 2.6: Equipment and innovation investment adoption rates across 34 consolidated provinces, ', base_year, ' and ', snapshot_year, '.')}
p_equipment_34 <- ggplot() +
  geom_sf(data = vn_adoption_34_map, aes(fill = equipment_adoption_rate),
          colour = "#6B7280", linewidth = 0.02) +
  geom_sf(data = island_polygons, inherit.aes = FALSE,
          fill = "#12355B", colour = "white", linewidth = 0.12) +
  geom_sf_text(data = island_labels, aes(label = island_group), inherit.aes = FALSE,
               size = 3, colour = "#12355B", fontface = "bold") +
  facet_wrap(~ survey_year, ncol = 2, drop = TRUE) +
  scale_fill_gradient(
    low = "#E6F5F2", high = "#006D77",
    labels = percent_format(accuracy = 1), na.value = "#F1F5F9",
    name = "Equipment\nadoption rate"
  ) +
  labs(
    title    = glue("{target_sector}: Equipment Investment Adoption By Province"),
    subtitle = "34-province map; share of firms with machinery and equipment investment, 2019 and 2023",
    caption  = "Adoption rate = firms with equipment investment / total firms in province-year."
  ) +
  coord_sf(xlim = c(102, 117), ylim = c(7, 24), expand = FALSE, datum = NA) +
  theme_void(base_size = 12) +
  theme(
    plot.title      = element_text(face = "bold", size = 16, colour = "#12355B"),
    plot.subtitle   = element_text(size = 10.5, colour = "#5B6F8A"),
    plot.caption    = element_text(size = 9, colour = "#6B7280", hjust = 0),
    strip.text      = element_text(face = "bold", size = 12, colour = "#12355B"),
    panel.spacing   = unit(0.8, "lines"),
    legend.position = "bottom",
    legend.title    = element_text(face = "bold"),
    legend.text     = element_text(size = 9),
    plot.margin     = margin(10, 10, 10, 10)
  )

p_equipment_34

p_innovation_34 <- ggplot() +
  geom_sf(data = vn_adoption_34_map, aes(fill = innovation_adoption_rate),
          colour = "#6B7280", linewidth = 0.03) +
  geom_sf(data = island_polygons, inherit.aes = FALSE,
          fill = "#12355B", colour = "white", linewidth = 0.15) +
  geom_sf_text(data = island_labels, aes(label = island_group), inherit.aes = FALSE,
               size = 3, colour = "#12355B", fontface = "bold") +
  facet_wrap(~ survey_year, ncol = 2, drop = TRUE) +
  scale_fill_gradient(
    low = "#FFFBF0", high = "#E76F51",
    labels = percent_format(accuracy = 1), na.value = "#F1F5F9",
    name = "Innovation\nadoption rate"
  ) +
  labs(
    title    = glue("{target_sector}: Innovation Investment Adoption By Province"),
    subtitle = "34-province map; share of firms with innovation investment, 2019 and 2023",
    caption  = "Adoption rate = firms with innovation investment / total firms in province-year."
  ) +
  coord_sf(xlim = c(102, 117), ylim = c(7, 24), expand = FALSE, datum = NA) +
  theme_void(base_size = 12) +
  theme(
    plot.title      = element_text(face = "bold", size = 16, colour = "#12355B"),
    plot.subtitle   = element_text(size = 10.5, colour = "#5B6F8A"),
    plot.caption    = element_text(size = 9, colour = "#6B7280", hjust = 0),
    strip.text      = element_text(face = "bold", size = 12, colour = "#12355B"),
    panel.spacing   = unit(0.8, "lines"),
    legend.position = "bottom",
    legend.title    = element_text(face = "bold"),
    legend.text     = element_text(size = 9),
    plot.margin     = margin(10, 10, 10, 10)
  )

p_innovation_34
```

```{r page2-visual5-interpretation, results='asis'}
top_equip_2023 <- province_adoption_34 %>%
  filter(survey_year == snapshot_year) %>%
  arrange(desc(equipment_adoption_rate)) %>%
  slice(1)
top_innov_2023 <- province_adoption_34 %>%
  filter(survey_year == snapshot_year) %>%
  arrange(desc(innovation_adoption_rate)) %>%
  slice(1)

equip_2019_row  <- province_adoption_34 %>%
  filter(survey_year == base_year, new_province_name == top_equip_2023$new_province_name)
equip_2019_rate <- if (nrow(equip_2019_row) > 0) equip_2019_row$equipment_adoption_rate[1] else NA_real_
equip_change_pp <- if (!is.na(equip_2019_rate)) {
  round((top_equip_2023$equipment_adoption_rate - equip_2019_rate) * 100, 1)
} else { NA_real_ }

equip_change_msg <- if_else(
  !is.na(equip_change_pp),
  glue(", {if_else(equip_change_pp >= 0, '+', '')}**{equip_change_pp} percentage points** change from {base_year}"),
  ""
)

# Innovation province change pp
innov_2019_row_map  <- province_adoption_34 %>%
  filter(survey_year == base_year, new_province_name == top_innov_2023$new_province_name)
innov_2019_rate_map <- if (nrow(innov_2019_row_map) > 0) innov_2019_row_map$innovation_adoption_rate[1] else NA_real_
innov_change_pp_map <- if (!is.na(innov_2019_rate_map)) {
  round((top_innov_2023$innovation_adoption_rate - innov_2019_rate_map) * 100, 1)
} else { NA_real_ }

innov_change_msg <- if_else(
  !is.na(innov_change_pp_map),
  glue(", {if_else(innov_change_pp_map >= 0, '+', '')}**{innov_change_pp_map} percentage points** change from {base_year}"),
  ""
)

# Concentration ratio: top province vs sector provincial average
national_equip_avg <- province_adoption_34 %>%
  filter(survey_year == snapshot_year) %>%
  summarise(v = mean(equipment_adoption_rate, na.rm = TRUE)) %>% pull(v)

concentration_ratio <- safe_divide(top_equip_2023$equipment_adoption_rate, national_equip_avg)

concentration_msg <- case_when(
  is.na(concentration_ratio) ~ "",
  concentration_ratio >= 2.0 ~
    glue("The leading province's equipment adoption rate is more than double the provincial average ({percent(national_equip_avg, accuracy = 0.1)}), indicating high geographic concentration in technology adoption."),
  concentration_ratio >= 1.3 ~
    glue("The leading province's rate is {round(concentration_ratio, 1)}x the provincial average ({percent(national_equip_avg, accuracy = 0.1)}), suggesting moderate concentration with meaningful room for regional diffusion."),
  TRUE ~
    glue("The leading province's rate is close to the provincial average ({percent(national_equip_avg, accuracy = 0.1)}), indicating relatively uniform adoption across provinces.")
)

# Same or different leading province for equipment vs innovation
same_province_msg <- if_else(
  !is.na(top_equip_2023$new_province_name) && !is.na(top_innov_2023$new_province_name) &&
    top_equip_2023$new_province_name == top_innov_2023$new_province_name,
  glue("Both investment types are led by the same province ({top_equip_2023$new_province_name}), suggesting a single technology hub dominates the sector's geography."),
  "The leading province for equipment adoption differs from that for innovation adoption, suggesting distinct regional ecosystems for hardware and capability investment."
)

interp(glue(
  "In {snapshot_year}, the highest equipment adoption rate was in **{top_equip_2023$new_province_name}** ",
  "at **{percent(top_equip_2023$equipment_adoption_rate, accuracy = 0.1)}**{equip_change_msg}. ",
  "The highest innovation adoption rate was in **{top_innov_2023$new_province_name}** ",
  "at **{percent(top_innov_2023$innovation_adoption_rate, accuracy = 0.1)}**{innov_change_msg}. ",
  "{concentration_msg} ",
  "{same_province_msg}"
))
```

::: part-divider
<span>Part III: Productivity Dynamics</span>
:::

# 3. Productivity Dynamics & Five-Driver Decomposition

```{r s3-prep, include=FALSE}
# ============================================================
# SECTION 3: DECOMPOSITION DATA PREP
# Robust version for LP_decomposition_CGEsector.xlsx
# Handles: SectorName / CGE_sectorname, numeric or character Year,
# trailing spaces, multiple sheets, and missing AVG rows.
# ============================================================

have_decomp_file <- !is.null(params$decomp_path) &&
  params$decomp_path != "" &&
  file.exists(params$decomp_path)

decomp_raw <- if (have_decomp_file) {
  readxl::excel_sheets(params$decomp_path) %>%
    set_names() %>%
    purrr::map_dfr(~ readxl::read_excel(params$decomp_path, sheet = .x), .id = "sheet_name")
} else {
  tibble()
}

clean_txt <- function(x) {
  x %>%
    as.character() %>%
    stringr::str_squish() %>%
    stringr::str_to_lower()
}

find_first_col <- function(data, candidates) {
  hit <- intersect(candidates, names(data))
  if (length(hit) == 0) NA_character_ else hit[1]
}

sector_col <- find_first_col(
  decomp_raw,
  c("CGE_sectorname", "CGE_sector", "SectorName", "sector", "Sector", "sector_name", "CGE sector name")
)

year_col <- find_first_col(
  decomp_raw,
  c("Year", "year", "survey_year", "Survey year")
)

required_decomp_cols <- c(
  "gKD", "g_tech_change", "g_effi_change",
  "g_equip_change", "g_inno_change",
  "gTFP", "gLP_macro", "gLP_check"
)

decomp_sector <- if (nrow(decomp_raw) > 0 && !is.na(sector_col) && !is.na(year_col)) {
  decomp_raw %>%
    mutate(
      sector_match_key = clean_txt(.data[[sector_col]]),
      target_match_key = clean_txt(target_sector)
    ) %>%
    filter(sector_match_key == target_match_key) %>%
    rename(Year = all_of(year_col)) %>%
    mutate(Year = suppressWarnings(as.integer(as.character(Year))))
} else {
  tibble()
}

if (nrow(decomp_sector) > 0) {
  for (cc in required_decomp_cols) {
    if (!cc %in% names(decomp_sector)) decomp_sector[[cc]] <- NA_real_
  }
  decomp_sector <- decomp_sector %>%
    mutate(across(all_of(required_decomp_cols), ~ suppressWarnings(as.numeric(.x))))
}

decomp_years <- if (nrow(decomp_sector) > 0 && "Year" %in% names(decomp_sector)) {
  decomp_sector %>%
    filter(Year %in% 2020:2023) %>%
    select(Year, all_of(required_decomp_cols)) %>%
    arrange(Year)
} else {
  tibble(
    Year = integer(),
    gKD = numeric(),
    g_tech_change = numeric(),
    g_effi_change = numeric(),
    g_equip_change = numeric(),
    g_inno_change = numeric(),
    gTFP = numeric(),
    gLP_macro = numeric(),
    gLP_check = numeric()
  )
}

decomp_avg <- if (nrow(decomp_years) > 0) {
  decomp_years %>%
    summarise(
      across(all_of(required_decomp_cols), ~ mean(.x, na.rm = TRUE)),
      .groups = "drop"
    ) %>%
    mutate(across(all_of(required_decomp_cols), ~ ifelse(is.nan(.x), NA_real_, .x)))
} else {
  tibble(
    gKD = NA_real_,
    g_tech_change = NA_real_,
    g_effi_change = NA_real_,
    g_equip_change = NA_real_,
    g_inno_change = NA_real_,
    gTFP = NA_real_,
    gLP_macro = NA_real_,
    gLP_check = NA_real_
  )
}

have_decomp <- nrow(decomp_avg) == 1 &&
  is.finite(decomp_avg$gLP_macro) &&
  !is.na(decomp_avg$gLP_macro)

# Diagnostic objects for quick checking in console if needed:
# decomp_sector_match_n should be 5 rows for one sector-year panel, usually 2019–2023.
decomp_sector_match_n <- nrow(decomp_sector)
decomp_years_match_n <- nrow(decomp_years)
decomp_available_sectors <- if (!is.na(sector_col) && nrow(decomp_raw) > 0) {
  decomp_raw %>% distinct(.data[[sector_col]]) %>% arrange(.data[[sector_col]])
} else tibble()

if (have_decomp) {
  drv_named <- tibble(
    key = c("KD", "TC", "CE", "ZE1", "ZE2"),
    name = c("Capital Deepening", "Technology Frontier Change", "Efficiency Change", "Hard-Tech Diffusion", "Soft-Tech Diffusion"),
    label = c("Capital
Deepening", "Technology
Frontier Change", "Efficiency
Change", "Hard-Tech
Diffusion", "Soft-Tech
Diffusion"),
    value = c(decomp_avg$gKD, decomp_avg$g_tech_change, decomp_avg$g_effi_change, decomp_avg$g_equip_change, decomp_avg$g_inno_change)
  ) %>%
    mutate(
      share     = value / decomp_avg$gLP_check,
      share_pct = share * 100
    )

  dominant_driver <- drv_named %>% arrange(desc(abs(value))) %>% slice(1)
  diffusion_total <- decomp_avg$g_equip_change + decomp_avg$g_inno_change
  diffusion_share  <- safe_divide(diffusion_total, decomp_avg$gLP_check)
  frontier_share   <- safe_divide(decomp_avg$g_tech_change, decomp_avg$gLP_check)
  capital_share    <- safe_divide(decomp_avg$gKD, decomp_avg$gLP_check)
  efficiency_share <- safe_divide(decomp_avg$g_effi_change, decomp_avg$gLP_check)

  regime_decomp <- if (is.na(decomp_avg$gLP_check) || decomp_avg$gLP_check <= 0) {
    "contraction"
  } else {
    case_when(
      capital_share > 0.50 ~ "capital-led",
      frontier_share > 0.40 ~ "frontier-led",
      diffusion_share > 0.40 ~ "diffusion-led",
      efficiency_share > 0.40 ~ "efficiency-led",
      TRUE ~ "mixed"
    )
  }
} else {
  drv_named <- tibble()
  dominant_driver <- tibble(name = "not available", value = NA_real_, share = NA_real_)
  diffusion_total <- diffusion_share <- frontier_share <- capital_share <- efficiency_share <- NA_real_
  regime_decomp <- "unavailable"
}
```


## 3.1 Average Driver Contributions

<p class="section-desc">
Waterfall decomposition of average annual labour productivity growth into five drivers. The chart identifies the dominant growth engine and the balance between frontier expansion and technology diffusion.
</p>

```{r s3-waterfall-data, include=FALSE}
if (have_decomp) {
  wf <- drv_named %>%
    mutate(
      idx = row_number(),
      x_start = idx - 0.4,
      x_end = idx + 0.4,
      cum_before = cumsum(lag(value, default = 0)),
      cum_after = cumsum(value),
      y_bottom = pmin(cum_before, cum_after),
      y_top = pmax(cum_before, cum_after),
      label_y = pmax(cum_before, cum_after) + 0.15,
      bar_label = paste0(ifelse(value >= 0, "+", ""), round(value, 1),
                         "\n(", round(share_pct, 0), "%)")
    )

  total_idx <- nrow(wf) + 1

  total_bar <- tibble(
    key = "TOTAL", name = "Total Labour Productivity Growth", label = "Total Labour\nProductivity Growth",
    value = decomp_avg$gLP_check, share_pct = 100, idx = total_idx,
    x_start = total_idx - 0.4, x_end = total_idx + 0.4,
    cum_before = 0, cum_after = decomp_avg$gLP_check,
    y_bottom = pmin(0, decomp_avg$gLP_check), y_top = pmax(0, decomp_avg$gLP_check),
    label_y = pmax(0, decomp_avg$gLP_check) + 0.15,
    bar_label = paste0(ifelse(decomp_avg$gLP_check >= 0, "+", ""),
                       round(decomp_avg$gLP_check, 1), "\n(100%)")
  )

  wf_all <- bind_rows(wf, total_bar)

  connectors <- wf %>%
    mutate(x = x_end, xend = lead(x_start, default = total_idx - 0.4), y = cum_after, yend = cum_after) %>%
    select(x, xend, y, yend)

  wf_axis_top <- max(c(wf_all$y_top, 0), na.rm = TRUE) * 1.25 + 0.5
  wf_axis_bot <- min(c(wf_all$y_bottom, 0), na.rm = TRUE) - 0.5
}
```

```{r s3-waterfall-plot, results='asis', fig.width=12, fig.height=6.5, fig.cap='Figure 3.1. Average driver contributions to annual labour productivity growth, 2020 to 2023. Each bar shows a driver contribution in percentage points; the right-hand bar is the total.'}
if (have_decomp) {
  driver_cols <- c("KD" = "#2A9D8F", "TC" = "#12355B", "CE" = "#4CAF50", "ZE1" = "#4A6FA5", "ZE2" = "#F4A261", "TOTAL" = "#12355B")

  ggplot() +
    geom_hline(yintercept = 0, colour = "#6B7280", linewidth = 0.6) +
    geom_segment(data = connectors, aes(x = x, xend = xend, y = y, yend = yend),
                 linetype = "dashed", colour = "#8898A7", linewidth = 0.55) +
    geom_rect(data = wf_all, aes(xmin = x_start, xmax = x_end, ymin = y_bottom, ymax = y_top, fill = key),
              colour = "white", linewidth = 0.6) +
    geom_text(data = wf_all,
              aes(x = idx, y = label_y,
                  label = bar_label,
                  colour = key),
              fontface = "bold", size = 4.2, lineheight = 0.92, vjust = 0) +
    scale_fill_manual(values = driver_cols) +
    scale_colour_manual(values = driver_cols) +
    scale_x_continuous(breaks = c(wf$idx, total_idx), labels = c(wf$label, "Total Labour\nProductivity Growth"),
                       expand = expansion(add = c(0.3, 0.3))) +
    scale_y_continuous(name = "Contribution to annual labour productivity growth (percentage points)",
                       limits = c(min(0, wf_axis_bot), wf_axis_top),
                       breaks = pretty(c(wf_axis_bot, wf_axis_top), n = 8),
                       expand = expansion(mult = c(0.02, 0.02))) +
    labs(title = glue("{target_sector}: Drivers of Annual Labour Productivity Growth ({base_year} to {snapshot_year})"),
         subtitle = "Average annual contribution to labour productivity growth (percentage points)",
         x = NULL) +
    theme_sector() +
    theme(legend.position = "none",
          axis.text.x = element_text(face = "bold", colour = "#12355B", size = 11, lineheight = 0.95),
          panel.grid.major.x = element_blank())
} else {
  cat("<div class='data-note'>Decomposition data for this sector is unavailable; the waterfall chart has been skipped.</div>")
}
```

```{r s3-waterfall-interpretation, results='asis'}
if (have_decomp) {
  diff_frontier_msg <- case_when(
    !is.na(diffusion_share) && !is.na(frontier_share) && diffusion_share > frontier_share + 0.10 ~
      glue("Diffusion-related drivers contributed **{percent(diffusion_share, accuracy = 1)}** of growth vs **{percent(frontier_share, accuracy = 1)}** from frontier change, meaning capability is spreading across the firm distribution faster than leading firms are advancing the boundary."),
    !is.na(diffusion_share) && !is.na(frontier_share) && frontier_share > diffusion_share + 0.10 ~
      glue("Technology Frontier Change contributed **{percent(frontier_share, accuracy = 1)}** of growth vs **{percent(diffusion_share, accuracy = 1)}** from diffusion, meaning leading firms are advancing faster than the broader sector can absorb , a pattern that tends to widen the frontier-laggard gap over time."),
    TRUE ~
      glue("Frontier change (**{percent(frontier_share, accuracy = 1)}**) and diffusion (**{percent(diffusion_share, accuracy = 1)}**) contributed in broadly similar magnitudes, indicating a relatively balanced productivity structure with neither channel dominating.")
  )

  regime_msg <- case_when(
    regime_decomp == "capital-led" ~
      "The sector is best characterised as **capital-led**: growth relies heavily on rising capital per worker and faces diminishing returns unless complemented by capability upgrading.",
    regime_decomp == "frontier-led" ~
      "The sector is best characterised as **frontier-led**: leading firms are carrying a large share of productivity improvement, creating risk of widening within-sector gaps if diffusion mechanisms do not keep pace.",
    regime_decomp == "diffusion-led" ~
      "The sector is best characterised as **diffusion-led**: adoption and capability spread are central to productivity growth, meaning gains are reaching a wider set of firms rather than concentrating at the frontier.",
    regime_decomp == "efficiency-led" ~
      "The sector is best characterised as **efficiency-led**: resource allocation and operational performance are the main growth channels, which are real but bounded ; once easy efficiency gains are captured, further growth requires a different engine.",
    regime_decomp == "mixed" ~
      "The sector has a **mixed** growth structure, with no single driver fully dominating, meaning no single policy lever in isolation is likely to materially shift the productivity trajectory.",
    TRUE ~ "The sector's growth regime could not be classified reliably from the available decomposition data."
  )

  interp(glue(
    "{target_sector} recorded **{round(decomp_avg$gLP_check, 2)} percentage points** of average annual labour productivity growth between **2020 and 2023**. ",
    "The largest contributor was **{dominant_driver$name}**, adding **{round(dominant_driver$value, 2)} percentage points** (**{percent(dominant_driver$share, accuracy = 1)}** of total growth). ",
    "Capital Deepening contributed **{round(decomp_avg$gKD, 2)} percentage points** and Total Factor Productivity contributed **{round(decomp_avg$gTFP, 2)} percentage points**.",
    "\n\n",
    "{diff_frontier_msg} {regime_msg} ",
    "This diagnosis matters for policy because the dominant driver identifies whether future productivity gains are more likely to come from capital accumulation, frontier innovation, diffusion mechanisms or efficiency upgrading."
  ))
}
```

## 3.2 Year-by-Year Driver Dynamics

<p class="section-desc">
Stacked decomposition of annual labour productivity growth. This visual shows whether the sector's growth structure remained stable or shifted over time.
</p>

```{r s3-stacked-plot, results='asis', fig.width=11, fig.height=6, fig.cap='Figure 3.2. Year-by-year decomposition of labour productivity growth, 2020 to 2023. Bars stack the five drivers; the red line tracks total labour productivity growth.'}
if (nrow(decomp_years) >= 1) {
  yearly <- decomp_years %>%
    filter(Year %in% 2020:2023) %>%
    select(Year, `Capital Deepening` = gKD, `Technology Frontier Change` = g_tech_change,
           `Efficiency Change` = g_effi_change, `Hard-Tech Diffusion` = g_equip_change,
           `Soft-Tech Diffusion` = g_inno_change, total = gLP_check) %>%
    pivot_longer(-c(Year, total), names_to = "Driver", values_to = "pp") %>%
    mutate(Driver = factor(Driver, levels = c("Capital Deepening", "Technology Frontier Change",
                                              "Efficiency Change", "Hard-Tech Diffusion", "Soft-Tech Diffusion")))

  driver_pal <- c("Capital Deepening" = "#2A9D8F", "Technology Frontier Change" = "#12355B",
                  "Efficiency Change" = "#4CAF50", "Hard-Tech Diffusion" = "#4A6FA5",
                  "Soft-Tech Diffusion" = "#F4A261")

  totals_df <- yearly %>% distinct(Year, total)

  ggplot(yearly, aes(x = factor(Year), y = pp, fill = Driver)) +
    geom_col(width = 0.7, colour = "white", linewidth = 0.3) +
    geom_hline(yintercept = 0, colour = "#6B7280", linewidth = 0.7) +
    geom_line(data = totals_df, aes(x = factor(Year), y = total, group = 1),
              inherit.aes = FALSE, colour = "#E63946", linewidth = 1.3) +
    geom_point(data = totals_df, aes(x = factor(Year), y = total),
               inherit.aes = FALSE, colour = "#E63946", size = 3.5) +
    geom_text(data = totals_df, aes(x = factor(Year), y = total,
                                    label = paste0(ifelse(total >= 0, "+", ""), round(total, 1), " pp")),
              inherit.aes = FALSE, vjust = -0.9, fontface = "bold", colour = "#E63946", size = 3.6) +
    scale_fill_manual(values = driver_pal) +
    scale_y_continuous(labels = number_format(accuracy = 0.1), expand = expansion(mult = c(0.05, 0.2))) +
    labs(title = glue("{target_sector}: Year-by-Year Driver Decomposition"),
         subtitle = "Stacked bars show each driver's contribution (percentage points). Red line: total labour productivity growth.",
         x = NULL, y = "Contribution (percentage points)", fill = "Driver") +
    theme_sector() +
    theme(legend.position = "bottom", legend.title = element_text(face = "bold"))
} else {
  cat("<div class='data-note'>Year-by-year decomposition data unavailable for this sector.</div>")
}
```

```{r s3-stacked-interpretation, results='asis'}
if (nrow(decomp_years) >= 1) {
  LP_best_year <- decomp_years$Year[which.max(decomp_years$gLP_check)]
  LP_worst_year <- decomp_years$Year[which.min(decomp_years$gLP_check)]
  LP_best_val <- max(decomp_years$gLP_check, na.rm = TRUE)
  LP_worst_val <- min(decomp_years$gLP_check, na.rm = TRUE)

  yr_largest <- function(col) {
    vals <- decomp_years[[col]]
    if (all(is.na(vals))) return("not available")
    decomp_years$Year[which.max(abs(vals))]
  }

  yr_KD <- yr_largest("gKD")
  yr_TC <- yr_largest("g_tech_change")
  yr_CE <- yr_largest("g_effi_change")
  yr_Z1 <- yr_largest("g_equip_change")
  yr_Z2 <- yr_largest("g_inno_change")

  diffusion_2020 <- decomp_years %>% filter(Year == 2020) %>%
    summarise(v = sum(g_equip_change, g_inno_change, na.rm = TRUE)) %>% pull(v)
  diffusion_2023 <- decomp_years %>% filter(Year == 2023) %>%
    summarise(v = sum(g_equip_change, g_inno_change, na.rm = TRUE)) %>% pull(v)

  diffusion_trend_msg <- case_when(
    is.na(diffusion_2020) | is.na(diffusion_2023) ~
      "The diffusion trend across the period could not be measured.",
    diffusion_2023 > diffusion_2020 + 0.5 ~
      glue("Combined diffusion contributions strengthened from **{round(diffusion_2020, 1)} percentage points** in 2020 to **{round(diffusion_2023, 1)} percentage points** in 2023, suggesting broader technology absorption as the period progressed."),
    diffusion_2023 < diffusion_2020 - 0.5 ~
      glue("Combined diffusion contributions declined from **{round(diffusion_2020, 1)} percentage points** in 2020 to **{round(diffusion_2023, 1)} percentage points** in 2023, indicating that adoption gains were concentrated in the early years."),
    TRUE ~
      glue("Diffusion contributions remained broadly stable across the period (from **{round(diffusion_2020, 1)} percentage points** in 2020 to **{round(diffusion_2023, 1)} percentage points** in 2023), indicating a consistent but not accelerating absorption capacity.")
  )

  lp_range <- LP_best_val - LP_worst_val

  volatility_msg <- case_when(
    lp_range > 5 ~
      "The range between the best and worst year exceeds 5 percentage points, indicating high year-to-year volatility in the sector's productivity trajectory.",
    lp_range > 2 ~
      "Moderate year-to-year variation in labour productivity growth suggests the sector's performance is sensitive to annual shocks rather than following a smooth structural trend.",
    TRUE ~
      "Relatively stable year-to-year variation indicates that the sector's growth structure is consistent rather than driven by year-specific shocks."
  )

  interp(glue(
    "Labour productivity growth was highest in **{LP_best_year}** at **{round(LP_best_val, 2)} percentage points** and lowest in **{LP_worst_year}** at **{round(LP_worst_val, 2)} percentage points**. ",
    "{volatility_msg} ",
    "Capital Deepening made its largest annual contribution in **{yr_KD}**, Technology Frontier Change in **{yr_TC}**, Efficiency Change in **{yr_CE}**, Hard-Tech Diffusion in **{yr_Z1}**, and Soft-Tech Diffusion in **{yr_Z2}**.",
    "\n\n",
    "{diffusion_trend_msg} ",
    "The year-by-year pattern matters because a stable and broad-based contribution from diffusion is usually more sustainable than growth concentrated in a single year or a single channel."
  ))
}
```

## 3.3 Conditional Productivity Frontier

<p class="section-desc">
Quantile-regression frontier comparing 2019 and 2023. The frontier shows the productivity level reached by top-performing firms at comparable capital intensity.
</p>

```{r s3-frontier-prep, include=FALSE}
TARGET_SECTOR <- target_sector
YEAR_EARLY <- 2019
YEAR_LATE <- 2023
TAU <- 0.90
TAU_LO <- 0.80
TAU_HI <- 0.99
DF_SPLINE <- 3
UNIT_LABEL <- "mn VND"
UNIT_DIV <- 1
OUTLIER_CAP <- 0.97

frontier_sector_df <- sector_df %>%
  filter(survey_year %in% c(YEAR_EARLY, YEAR_LATE),
         !is.na(revenue_per_worker), !is.na(capital_intensity),
         revenue_per_worker > 0, capital_intensity > 0) %>%
  mutate(y_pw = revenue_per_worker / UNIT_DIV,
         k_pw = capital_intensity / UNIT_DIV) %>%
  group_by(survey_year) %>%
  filter(y_pw <= quantile(y_pw, OUTLIER_CAP, na.rm = TRUE),
         k_pw <= quantile(k_pw, OUTLIER_CAP, na.rm = TRUE)) %>%
  ungroup()

n_early <- sum(frontier_sector_df$survey_year == YEAR_EARLY)
n_late <- sum(frontier_sector_df$survey_year == YEAR_LATE)

fit_one_quantile <- function(yr_data, tau, df_spline, n_grid) {
  fit <- quantreg::rq(y_pw ~ splines::ns(k_pw, df = df_spline), tau = tau, data = yr_data)
  k_grid <- seq(min(yr_data$k_pw, na.rm = TRUE), max(yr_data$k_pw, na.rm = TRUE), length.out = n_grid)
  tibble(k_pw = k_grid, y = as.numeric(predict(fit, newdata = data.frame(k_pw = k_grid))))
}

fit_frontier <- function(yr_data, tau, df_spline = 5, n_grid = 200) {
  main <- fit_one_quantile(yr_data, tau, df_spline, n_grid)
  lo <- fit_one_quantile(yr_data, TAU_LO, df_spline, n_grid)
  hi <- fit_one_quantile(yr_data, TAU_HI, df_spline, n_grid)
  tibble(k_pw = main$k_pw, fit = main$y, lower = lo$y, upper = hi$y)
}

front_early <- frontier_sector_df %>% filter(survey_year == YEAR_EARLY) %>%
  fit_frontier(TAU, DF_SPLINE) %>% mutate(year = YEAR_EARLY)
front_late <- frontier_sector_df %>% filter(survey_year == YEAR_LATE) %>%
  fit_frontier(TAU, DF_SPLINE) %>% mutate(year = YEAR_LATE)

avg_pts <- frontier_sector_df %>%
  group_by(survey_year) %>%
  summarise(
    k_avg = if ("labour_headcount_end" %in% names(frontier_sector_df)) weighted.mean(k_pw, w = labour_headcount_end, na.rm = TRUE) else mean(k_pw, na.rm = TRUE),
    y_avg = if ("labour_headcount_end" %in% names(frontier_sector_df)) weighted.mean(y_pw, w = labour_headcount_end, na.rm = TRUE) else mean(y_pw, na.rm = TRUE),
    .groups = "drop"
  )

COL_EARLY <- "#12355B"
COL_LATE <- "#2A9D8F"
label_early_f <- glue("Frontier {YEAR_EARLY}")
label_late_f <- glue("Frontier {YEAR_LATE}")
label_early_av <- glue("Industry average {YEAR_EARLY}")
label_late_av <- glue("Industry average {YEAR_LATE}")

colour_map <- c(setNames(COL_EARLY, label_early_f), setNames(COL_LATE, label_late_f),
                setNames(COL_EARLY, label_early_av), setNames(COL_LATE, label_late_av))
linetype_map <- c(setNames("solid", label_early_f), setNames("solid", label_late_f),
                  setNames("blank", label_early_av), setNames("blank", label_late_av))
shape_map <- c(setNames(NA, label_early_f), setNames(NA, label_late_f),
               setNames(16L, label_early_av), setNames(16L, label_late_av))
```

```{r s3-frontier-plot, fig.width=11, fig.height=6.5, fig.cap='Figure 3.3. Conditional productivity frontier, 2019 and 2023. The frontier line is the 90th-percentile quantile regression with a natural spline; the shaded band spans the 80th to 99th percentile frontier zone.'}
if (n_early >= 30 && n_late >= 30) {
  ggplot() +
    geom_ribbon(data = front_late, aes(x = k_pw, ymin = lower, ymax = upper), fill = "grey80", alpha = 0.55) +
    geom_ribbon(data = front_early, aes(x = k_pw, ymin = lower, ymax = upper), fill = "grey60", alpha = 0.45) +
    geom_line(data = front_early, aes(x = k_pw, y = fit, colour = label_early_f, linetype = label_early_f), linewidth = 1.0) +
    geom_line(data = front_late, aes(x = k_pw, y = fit, colour = label_late_f, linetype = label_late_f), linewidth = 1.0) +
    geom_point(data = avg_pts %>% filter(survey_year == YEAR_EARLY),
               aes(x = k_avg, y = y_avg, colour = label_early_av, shape = label_early_av), size = 3.5) +
    geom_point(data = avg_pts %>% filter(survey_year == YEAR_LATE),
               aes(x = k_avg, y = y_avg, colour = label_late_av, shape = label_late_av), size = 3.5) +
    scale_colour_manual(values = colour_map, breaks = c(label_late_f, label_early_f, label_late_av, label_early_av)) +
    scale_linetype_manual(values = linetype_map, breaks = c(label_late_f, label_early_f, label_late_av, label_early_av)) +
    scale_shape_manual(values = shape_map, breaks = c(label_late_f, label_early_f, label_late_av, label_early_av)) +
    guides(colour = guide_legend(override.aes = list(linetype = c("solid", "solid", "blank", "blank"),
                                                      shape = c(NA, NA, 16, 16), size = c(1, 1, 3, 3))),
           linetype = "none", shape = "none") +
    labs(
      title = glue("Conditional frontier for {TARGET_SECTOR} {YEAR_EARLY}–{YEAR_LATE}"),
      x = glue("Real capital per worker ({UNIT_LABEL}, 2019 prices)"),
      y = glue("Real output per worker ({UNIT_LABEL}, 2019 prices)"),
      colour = NULL,
      caption = glue("Note: Frontier line is the {round(TAU * 100)}th-percentile quantile regression with natural spline ",
                     "(df = {DF_SPLINE}). Shaded band spans the {round(TAU_LO * 100)}th–{round(TAU_HI * 100)}th percentile ",
                     "frontier zone. Industry average is labour-weighted mean. Source: GSO Enterprise Survey, 2019–2023.")
    ) +
    theme_sector() +
    theme(plot.caption    = element_text(size = 9, colour = "#6B7280", hjust = 0, margin = margin(t = 10)),
          legend.position  = "bottom",
          legend.text      = element_text(size = 10),
          legend.key.width = unit(1.8, "cm"))
} else {
  cat("<div class='data-note'>Conditional frontier data unavailable or insufficient for this sector (fewer than 30 firms in one or both years).</div>")
}
```

```{r s3-frontier-interpretation, results='asis'}
if (n_early >= 30 && n_late >= 30) {
  avg_early <- avg_pts %>% filter(survey_year == YEAR_EARLY)
  avg_late <- avg_pts %>% filter(survey_year == YEAR_LATE)

  avg_lp_change <- safe_divide(avg_late$y_avg[1], avg_early$y_avg[1]) - 1
  avg_cap_change <- safe_divide(avg_late$k_avg[1], avg_early$k_avg[1]) - 1

  frontier_mid_early <- front_early %>%
    filter(k_pw >= quantile(k_pw, 0.45, na.rm = TRUE), k_pw <= quantile(k_pw, 0.55, na.rm = TRUE)) %>%
    summarise(v = mean(fit, na.rm = TRUE)) %>% pull(v)

  frontier_mid_late <- front_late %>%
    filter(k_pw >= quantile(k_pw, 0.45, na.rm = TRUE), k_pw <= quantile(k_pw, 0.55, na.rm = TRUE)) %>%
    summarise(v = mean(fit, na.rm = TRUE)) %>% pull(v)

  frontier_mid_change <- safe_divide(frontier_mid_late, frontier_mid_early) - 1

  frontier_msg <- if_else(!is.na(frontier_mid_change) && frontier_mid_change > 0,
                          glue("The conditional frontier shifted upward by approximately **{percent(frontier_mid_change, accuracy = 0.1)}** around the mid-range of capital intensity."),
                          "The conditional frontier did not clearly shift upward around the mid-range of capital intensity.")

  avg_msg <- if_else(!is.na(avg_lp_change) && avg_lp_change > 0,
                     glue("The labour-weighted industry average output per worker also increased by **{percent(avg_lp_change, accuracy = 0.1)}**, while average capital per worker changed by **{percent(avg_cap_change, accuracy = 0.1)}**."),
                     glue("The labour-weighted industry average output per worker did not materially improve, despite a **{percent(avg_cap_change, accuracy = 0.1)}** change in average capital per worker."))

  structural_interp_msg <- case_when(
    is.na(frontier_mid_change) ~
      "The frontier shift cannot be assessed from available data.",
    frontier_mid_change > 0.10 & !is.na(avg_lp_change) & avg_lp_change > 0 ~
      glue("Both the frontier and the industry average improved. If the frontier outpaced the average (which is consistent with a wide and stable leader-laggard gap) it signals that diffusion is not fully keeping pace with frontier advancement."),
    frontier_mid_change > 0.10 & (is.na(avg_lp_change) | avg_lp_change <= 0) ~
      "The frontier shifted upward but the industry average did not improve materially : the clearest signal of a diffusion constraint: leading firms are advancing while most firms are not.",
    frontier_mid_change <= 0 & !is.na(avg_lp_change) & avg_lp_change > 0 ~
      "The industry average improved despite a flat or declining frontier, consistent with catch-up by laggard firms rather than frontier-driven growth.",
    TRUE ~
      "Neither the frontier nor the industry average showed clear improvement, suggesting the sector's productivity structure was broadly stagnant over the period."
  )

  interp(glue(
    "{frontier_msg} ",
    "{avg_msg} ",
    "{structural_interp_msg} ",
    "Structurally, the chart compares frontier movement with the average firm, showing whether productivity gains are concentrated at the top or absorbing more broadly. ",
    "A persistent distance between the frontier and the average is consistent with a capability diffusion gap rather than a simple lack of capital."
  ))
}
```

::: part-divider
<span>Part IV: Summary and Policy Insights</span>
:::

# 4. Summary Highlights & Policy Insights

## 4.1 Summary Highlights

<p class="section-desc">
Concise synthesis of the sector structure, technology adoption and productivity decomposition results. Each block reports a key numerical signal and its structural interpretation.
</p>

```{r s4-prep, include=FALSE}
economy_snapshot <- all_sector_year %>%
  filter(survey_year == snapshot_year) %>%
  summarise(total_revenue = sum(revenue, na.rm = TRUE), total_labour = sum(labour, na.rm = TRUE), .groups = "drop")

sector_snapshot <- sector_year %>%
  filter(survey_year == snapshot_year) %>%
  summarise(revenue = sum(revenue, na.rm = TRUE), labour = sum(labour, na.rm = TRUE), .groups = "drop")

rev_share_econ <- safe_divide(sector_snapshot$revenue, economy_snapshot$total_revenue)
lab_share_econ <- safe_divide(sector_snapshot$labour, economy_snapshot$total_labour)

sector_ranks_snapshot <- all_sector_year %>%
  filter(survey_year == snapshot_year) %>%
  mutate(rank_revenue = min_rank(desc(revenue)), rank_labour = min_rank(desc(labour))) %>%
  filter(CGE_sectorname == target_sector)

ownership_summary_5yr <- sector_df %>%
  filter(survey_year %in% analysis_years, !is.na(firm_ownership_type_factor)) %>%
  group_by(firm_ownership_type_factor) %>%
  summarise(revenue = sum(total_net_revenue, na.rm = TRUE), .groups = "drop") %>%
  mutate(revenue_share = safe_divide(revenue, sum(revenue, na.rm = TRUE))) %>%
  arrange(desc(revenue_share))

dominant_owner <- if (nrow(ownership_summary_5yr) > 0) as.character(ownership_summary_5yr$firm_ownership_type_factor[1]) else "not available"
dominant_owner_share <- if (nrow(ownership_summary_5yr) > 0) ownership_summary_5yr$revenue_share[1] else NA_real_

adoption_snapshot <- sector_year %>%
  filter(survey_year == snapshot_year) %>%
  summarise(equip_rate = mean(equipment_adoption_rate, na.rm = TRUE),
            innov_rate = mean(innovation_adoption_rate, na.rm = TRUE),
            total_investment = sum(total_investment, na.rm = TRUE),
            innovation_investment = sum(innovation_investment, na.rm = TRUE),
            .groups = "drop")

adoption_base <- sector_year %>%
  filter(survey_year == base_year) %>%
  summarise(equip_rate_base = mean(equipment_adoption_rate, na.rm = TRUE),
            innov_rate_base = mean(innovation_adoption_rate, na.rm = TRUE),
            .groups = "drop")

equip_rate_change_pp <- (adoption_snapshot$equip_rate - adoption_base$equip_rate_base) * 100
innov_rate_change_pp <- (adoption_snapshot$innov_rate - adoption_base$innov_rate_base) * 100
innovation_investment_share <- safe_divide(adoption_snapshot$innovation_investment, adoption_snapshot$total_investment)

frontier_gap_latest <- if (exists("frontier_year_metrics")) {
  tmp <- frontier_year_metrics %>% filter(survey_year == snapshot_year, Metric == "Labour productivity")
  if (nrow(tmp) > 0) tmp$Ratio[1] else NA_real_
} else {
  sector_df %>%
    mutate(firm_productivity = safe_divide(total_net_revenue, labour_headcount_end)) %>%
    filter(survey_year == snapshot_year, labour_headcount_end > 0, firm_productivity > 0) %>%
    mutate(productivity_rank = percent_rank(firm_productivity)) %>%
    summarise(top = median(firm_productivity[productivity_rank >= 0.97], na.rm = TRUE),
              rest = median(firm_productivity[productivity_rank < 0.97], na.rm = TRUE),
              ratio = safe_divide(top, rest), .groups = "drop") %>%
    pull(ratio)
}

top_equip_province <- if (exists("province_adoption_34")) {
  province_adoption_34 %>% filter(survey_year == snapshot_year) %>%
    arrange(desc(equipment_adoption_rate)) %>% slice(1)
} else tibble(new_province_name = "not available", equipment_adoption_rate = NA_real_)

top_innov_province <- if (exists("province_adoption_34")) {
  province_adoption_34 %>% filter(survey_year == snapshot_year) %>%
    arrange(desc(innovation_adoption_rate)) %>% slice(1)
} else tibble(new_province_name = "not available", innovation_adoption_rate = NA_real_)

lp_growth_avg <- if (have_decomp) decomp_avg$gLP_check else NA_real_
dominant_name <- if (have_decomp) dominant_driver$name else "not available"
dominant_value <- if (have_decomp) dominant_driver$value else NA_real_
dominant_share <- if (have_decomp) dominant_driver$share else NA_real_

# ── C1: Frontier gap trend ────────────────────────────────────────────────────
frontier_gap_base <- tryCatch({
  sector_df %>%
    mutate(firm_productivity = safe_divide(total_net_revenue, labour_headcount_end)) %>%
    filter(survey_year == base_year, labour_headcount_end > 0, firm_productivity > 0) %>%
    mutate(productivity_rank = percent_rank(firm_productivity)) %>%
    summarise(top  = median(firm_productivity[productivity_rank >= 0.97], na.rm = TRUE),
              rest = median(firm_productivity[productivity_rank <  0.97], na.rm = TRUE),
              ratio = safe_divide(top, rest), .groups = "drop") %>%
    pull(ratio)
}, error = function(e) NA_real_)

frontier_trend_msg <- case_when(
  is.na(frontier_gap_base) | is.na(frontier_gap_latest) ~
    "not available across both years",
  frontier_gap_latest > frontier_gap_base * 1.15 ~
    glue("widened from **{round(frontier_gap_base, 1)}x** in {base_year} to **{round(frontier_gap_latest, 1)}x** in {snapshot_year} , indicating that diffusion from leading firms to the wider sector has not kept pace with frontier advancement"),
  frontier_gap_latest < frontier_gap_base * 0.85 ~
    glue("narrowed from **{round(frontier_gap_base, 1)}x** in {base_year} to **{round(frontier_gap_latest, 1)}x** in {snapshot_year} , indicating that laggard firms are catching up and diffusion mechanisms are working"),
  TRUE ~
    glue("remained broadly stable between **{round(frontier_gap_base, 1)}x** in {base_year} and **{round(frontier_gap_latest, 1)}x** in {snapshot_year} , suggesting the frontier-laggard structure is entrenched rather than converging")
)

# ── C2: HD / SD / efficiency split ────────────────────────────────────────────
hd_value   <- if (have_decomp) decomp_avg$g_equip_change else NA_real_
sd_value   <- if (have_decomp) decomp_avg$g_inno_change  else NA_real_
effi_value <- if (have_decomp) decomp_avg$g_effi_change  else NA_real_

diffusion_type_msg <- case_when(
  is.na(hd_value) | is.na(sd_value) ~ "",
  hd_value > sd_value + 0.5 ~
    "Within adoption, equipment investment (hard-tech diffusion) was the larger channel , pointing to capital-embodied upgrading that may not yet be matched by organisational and capability development.",
  sd_value > hd_value + 0.5 ~
    "Within adoption, innovation investment (soft-tech diffusion) was the larger channel , a relatively advanced pattern where capability and management practices are outpacing physical equipment as productivity drivers.",
  TRUE ~
    "Hard-tech (equipment) and soft-tech (innovation) diffusion contributed in roughly similar proportions, indicating a balanced adoption profile."
)

effi_sign_msg <- case_when(
  is.na(effi_value)  ~ "",
  effi_value < -0.5  ~
    glue("Efficiency change was negative (**{round(effi_value, 2)} percentage points**): laggard firms are falling further behind best-practice operations, acting as a drag on aggregate productivity that management and organisational improvement could address."),
  effi_value >  0.5  ~
    glue("Efficiency change was positive (**{round(effi_value, 2)} percentage points**): laggard firms are closing the gap to frontier operating standards, a healthy sign that soft capabilities are diffusing alongside investment."),
  TRUE               ~
    glue("Efficiency change was near zero (**{round(effi_value, 2)} percentage points**): laggard firm performance relative to best practice has not materially improved or deteriorated.")
)
```

```{r s4-highlights, results='asis'}
highlight_block <- function(title, body) {
  body <- gsub("\\*\\*(.*?)\\*\\*", "<strong>\\1</strong>", body, perl = TRUE)
  paste0('<div class="highlight-block"><span class="highlight-title">', title,
         '</span><div class="highlight-body">', body, '</div></div>')
}

# ── Safe display strings for highlight blocks (NA-proof) ─────────────────────
frontier_gap_highlight_str <- if (!is.na(frontier_gap_latest)) glue("{round(frontier_gap_latest, 1)}x") else "not available"
lp_growth_str       <- if (!is.na(lp_growth_avg))   glue("{round(lp_growth_avg, 2)} percentage points") else "not available"
dom_value_str       <- if (!is.na(dominant_value))  glue("{round(dominant_value, 2)} percentage points") else "not available"
dom_share_str       <- if (!is.na(dominant_share))  percent(dominant_share, accuracy = 1) else "not available"
dom_owner_share_str <- if (!is.na(dominant_owner_share)) percent(dominant_owner_share, accuracy = 0.1) else "not available"

structure_blocks <- c(
  highlight_block("Sector scale and labour footprint",
    glue("{target_sector} contributed **{percent(rev_share_econ, accuracy = 0.1)}** of all-sector revenue and **{percent(lab_share_econ, accuracy = 0.1)}** of employment in {snapshot_year}. It ranked **{ifelse(nrow(sector_ranks_snapshot) == 1, sector_ranks_snapshot$rank_revenue, 'not available')} of 34** by revenue and **{ifelse(nrow(sector_ranks_snapshot) == 1, sector_ranks_snapshot$rank_labour, 'not available')} of 34** by labour, indicating its relative economic scale.")),
  highlight_block("Ownership structure",
    glue("Across 2019–2023, **{dominant_owner}** accounted for **{dom_owner_share_str}** of pooled sector revenue. This ownership profile indicates whether sector growth is primarily domestic-enterprise-led, state-linked or foreign-investment-led.")),
  highlight_block("Investment and adoption diffusion",
    glue("By {snapshot_year}, equipment adoption reached **{percent(adoption_snapshot$equip_rate, accuracy = 0.1)}** of firms and innovation adoption reached **{percent(adoption_snapshot$innov_rate, accuracy = 0.1)}**. Compared with {base_year}, these changed by **{round(equip_rate_change_pp, 1)} percentage points** and **{round(innov_rate_change_pp, 1)} percentage points**, respectively.")),
  highlight_block("Frontier and laggard gap",
    glue("Leading firms were approximately **{frontier_gap_highlight_str}** more productive than laggards in {snapshot_year}. A large ratio points to a capability gap, while a narrowing ratio would suggest stronger diffusion from frontier firms to the wider sector.")),
  highlight_block("Productivity growth engine",
    glue("Average annual labour productivity growth was **{lp_growth_str}** over 2020–2023. The dominant driver was **{dominant_name}**, contributing **{dom_value_str}** (**{dom_share_str}** of total). {diffusion_type_msg} {effi_sign_msg}")),
  highlight_block("Regional technology concentration",
    glue("The highest equipment adoption rate in {snapshot_year} was observed in **{top_equip_province$new_province_name}** at **{percent(top_equip_province$equipment_adoption_rate, accuracy = 0.1)}**, while the highest innovation adoption rate was in **{top_innov_province$new_province_name}** at **{percent(top_innov_province$innovation_adoption_rate, accuracy = 0.1)}**.")),
  highlight_block("Frontier gap trend",
    glue("The productivity gap between leading and laggard firms has {frontier_trend_msg}. Whether this gap is narrowing over time is the central diagnostic for whether technology adoption is generating broad-based gains or concentrating productivity at the frontier."))
)

cat('<div class="highlight-grid">')
cat(paste(structure_blocks, collapse = "\n"))
cat('</div>')
```

## 4.2 Policy Implications

<p class="section-desc">
Sector-specific policy diagnosis based on the growth regime, frontier dynamics, diffusion quality, investment mix, ownership structure and regional adoption pattern.
</p>

```{r s4-policy, results='asis'}
diffusion_constraint <- have_decomp && !is.na(diffusion_share) && !is.na(frontier_gap_latest) &&
  frontier_gap_latest > 5 && diffusion_share < 0.30

innovation_constraint <- !is.na(innovation_investment_share) && innovation_investment_share < 0.20 &&
  !is.na(adoption_snapshot$innov_rate) && adoption_snapshot$innov_rate < 0.20

regional_constraint <- !is.na(top_equip_province$equipment_adoption_rate) &&
  !is.na(adoption_snapshot$equip_rate) &&
  top_equip_province$equipment_adoption_rate > adoption_snapshot$equip_rate * 1.5

main_constraint <- case_when(
  diffusion_constraint ~ "diffusion bottleneck",
  innovation_constraint ~ "innovation capability constraint",
  regime_decomp == "capital-led" ~ "capital-heavy growth with capability absorption risk",
  regime_decomp == "frontier-led" ~ "frontier-led growth with spillover risk",
  regime_decomp == "diffusion-led" ~ "diffusion-led growth that requires scaling support",
  regime_decomp == "efficiency-led" ~ "efficiency-led growth with management and allocation constraints",
  regional_constraint ~ "regional diffusion gap",
  TRUE ~ "mixed productivity constraint"
)

policy_priority <- case_when(
  main_constraint == "diffusion bottleneck" ~
    "Policy should prioritise SME technology absorption, supplier-development programmes and technical extension services that help non-frontier firms adopt existing technologies more effectively.",
  main_constraint == "innovation capability constraint" ~
    "Policy should strengthen innovation capability through applied R&D partnerships, digital capability programmes and targeted support for intangible investment rather than relying only on machinery upgrading.",
  main_constraint == "capital-heavy growth with capability absorption risk" ~
    "Policy should avoid relying solely on further capital expansion. The priority is complementary capability building, including workforce skills, management upgrading and operational know-how that raise returns to capital.",
  main_constraint == "frontier-led growth with spillover risk" ~
    "Policy should pair frontier innovation support with mechanisms that transfer capability outward, including supplier linkages, standards, training platforms and demonstration projects for lagging firms.",
  main_constraint == "diffusion-led growth that requires scaling support" ~
    "Policy should reinforce the existing diffusion channel by expanding access to investment finance, digital adoption support and regional technical assistance for smaller firms.",
  main_constraint == "efficiency-led growth with management and allocation constraints" ~
    "Policy should focus on management modernisation, competition pressure and resource-allocation improvements that help firms convert existing inputs into higher output.",
  main_constraint == "regional diffusion gap" ~
    "Policy should address regional unevenness by targeting provinces where technology adoption remains low, while using leading provinces as demonstration hubs for capability diffusion.",
  TRUE ~
    "Policy should use a balanced package of capability building, diffusion support and selective frontier upgrading, because no single constraint dominates the sector's productivity pattern."
)

# ── C3: Creation vs adoption gap framing ──────────────────────────────────────
creation_share <- if (have_decomp && !is.na(decomp_avg$g_tech_change) && !is.na(decomp_avg$gLP_check))
  safe_divide(decomp_avg$g_tech_change, decomp_avg$gLP_check) else NA_real_

adoption_gap_framing <- case_when(
  is.na(creation_share) | is.na(diffusion_share) ~ "",
  creation_share < 0.10 & diffusion_share < 0.30 ~
    " Both frontier advancement and adoption diffusion are constrained : the sector faces a dual gap where neither the creation of new technological capacity nor the spread of existing practices is generating strong momentum.",
  creation_share < 0.10 & diffusion_share >= 0.30 ~
    " Diffusion is the primary productivity engine but frontier creation is limited. Gains depend on spreading existing technology rather than generating new capability , sustainable in the short run but exposed to exhaustion of the adoptable stock.",
  creation_share >= 0.40 & diffusion_share < 0.20 ~
    " A frontier-diffusion imbalance: leading firms are advancing but the broader sector cannot absorb these advances. The bottleneck is diffusion capacity (absorption capability, skills and finance), not the supply of technology.",
  TRUE ~ ""
)

# ── C3: FDI-spillover risk flag ────────────────────────────────────────────────
fdi_spillover_risk <- isTRUE(
  !is.na(dominant_owner_share) &&
  dominant_owner == "FDI" &&
  dominant_owner_share > 0.50 &&
  regime_decomp == "capital-led"
)

ownership_spillover_msg <- if (fdi_spillover_risk) {
  glue(" The sector's FDI-dominated revenue structure ({percent(dominant_owner_share, accuracy = 0.1)} from foreign-invested firms) combined with a capital-led growth regime signals a passive FDI dependence risk: capital accumulation accelerates through foreign investment but domestic firm capability may not be advancing in parallel. Policy should prioritise domestic absorptive capacity, including supplier-linkage programmes, workforce capability platforms and technology extension, alongside any further FDI attraction efforts.")
} else {
  ""
}

# ── Safe display strings for policy box (NA-proof) ───────────────────────────
regime_decomp_str      <- if (have_decomp) regime_decomp else "not available (no decomposition data)"
frontier_gap_str       <- if (!is.na(frontier_gap_latest)) glue("{round(frontier_gap_latest, 1)}x") else "not available"
diffusion_share_str    <- if (!is.na(diffusion_share)) percent(diffusion_share, accuracy = 1) else "not available"
effi_str               <- if (!is.na(effi_value)) glue("{round(effi_value, 2)} percentage points") else "not available"
innov_invest_str       <- if (!is.na(innovation_investment_share)) percent(innovation_investment_share, accuracy = 0.1) else "not available"

cat(glue("
<div class='policy-box'>
<strong>Policy diagnosis.</strong> The evidence points to a <strong>{main_constraint}</strong> in {target_sector}. This diagnosis draws on the sector's <strong>{regime_decomp_str}</strong> decomposition profile, a frontier-laggard ratio of <strong>{frontier_gap_str}</strong>, diffusion's contribution of <strong>{diffusion_share_str}</strong> to growth, efficiency change of <strong>{effi_str}</strong>, and an innovation-investment share of <strong>{innov_invest_str}</strong> in {snapshot_year}.{adoption_gap_framing}
</div>

<div class='policy-box'>
<strong>Immediate policy implication.</strong> {policy_priority} The goal is not simply to raise investment volume, but to improve the mechanisms through which firms convert capital, equipment, innovation and management practices into sustained productivity gains.{ownership_spillover_msg}
</div>

<div class='policy-box'>
<strong>Implementation focus.</strong> For this sector, interventions should be sequenced around firm capability, technology absorption and diffusion quality. Where the frontier gap is widening, priority goes to mechanisms that transfer capability outward (supplier development, training platforms, standards and demonstration projects) rather than expanding investment volume alone. Where efficiency change is negative, management modernisation and operational upgrading offer the fastest near-term productivity dividend without requiring additional capital.
</div>
"))
```

## 4.3 Strategic Closing Statement

```{r s4-closing, results='asis'}
closing_statement <- case_when(
  regime_decomp == "capital-led" ~
    glue("Sustained productivity growth in {target_sector} will require a transition from capital accumulation toward capability upgrading. The sector's capital-led profile suggests diminishing returns to further physical investment unless complemented by improvements in workforce skills, management practice and technology absorption. The long-term competitiveness question is not how much capital is added, but how effectively that capital is converted into output by a wider range of firms. At Vietnam's current development stage, capital-led sectors that do not generate strong domestic capability diffusion risk becoming sources of capital efficiency decline rather than productivity leadership , particularly where FDI integration is high but domestic value-added and supplier linkages remain shallow."),
  regime_decomp == "frontier-led" ~
    glue("Sustained productivity growth in {target_sector} depends on whether frontier firms can act as engines of broader capability diffusion. The frontier-led growth pattern indicates that leading firms are advancing rapidly, but the gap between these firms and the wider sector risks becoming self-reinforcing unless mechanisms for technology transfer, supplier development and workforce mobility are strengthened. Long-term competitiveness requires that frontier gains eventually reach a broader base of producers. Frontier-led sectors carry high potential as national diffusion sources: when leading firms are connected to domestic supply chains, their technology decisions and workforce practices create spillover pathways that policy can reinforce through standards, linkage programmes and technology extension."),
  regime_decomp == "diffusion-led" ~
    glue("Sustained productivity growth in {target_sector} is well-served by its diffusion-led profile, in which capability spreads across the firm distribution rather than concentrating at the frontier. The key long-term risk is that this diffusion channel loses momentum if the stock of adoptable practices becomes exhausted and is not renewed by fresh frontier innovation. Maintaining competitiveness over the longer term will require pairing diffusion support with investments in new technology development. Diffusion-led sectors contribute directly to inclusive growth by spreading productivity gains beyond a narrow frontier ; these sectors are also the most exposed to technology exhaustion if the practices being diffused are not continuously renewed by frontier advancement or international technology transfer."),
  regime_decomp == "efficiency-led" ~
    glue("Sustained productivity growth in {target_sector} has drawn heavily on operational and managerial catch-up. Efficiency-led gains are real but bounded: once the most tractable improvements in resource allocation and management practice have been captured, further growth requires a different engine. The sector's long-term competitiveness depends on whether efficiency gains can be complemented by rising technology intensity and innovation capability. Efficiency-led gains are particularly important for employment quality: operational catch-up tends to improve conditions for the median worker, and sectors on this trajectory offer a path to productivity improvement that is inclusive before it becomes innovation-intensive."),
  regime_decomp == "contraction" ~
    glue("Restoring productivity growth in {target_sector} requires first diagnosing whether the recorded contraction reflects a cyclical shock or a structural shift. Standard productivity levers (diffusion support, capital upgrading, capability programmes) should be sequenced behind that diagnosis, since applying growth-mode policies to a structurally contracting sector risks misallocating resources toward activities that do not address the underlying cause. A priority in this regime is to preserve firm capability and workforce skills during the contraction phase: the absorptive capacity of surviving firms determines how quickly the sector can benefit from new technology investment when conditions stabilise."),
  regime_decomp == "unavailable" ~
    glue("Labour productivity decomposition data is not available for {target_sector} in this dataset. A structural assessment of the sector's growth regime cannot be made without the LP decomposition workbook covering this sector. Policy analysis should treat the structural and diffusion findings on pages 1 and 2 as the primary evidence base until decomposition data becomes available."),
  TRUE ~
    glue("Sustained productivity growth in {target_sector} will depend less on any single driver and more on the quality of interactions between capital accumulation, capability diffusion and operational efficiency. Over the longer term, the sector's competitiveness will increasingly depend on its ability to translate improvements across multiple channels into broad-based capability upgrading, workforce adaptation and more inclusive productivity gains. Sectors with mixed productivity regimes are typically at a structural transition point (neither clearly behind the diffusion frontier nor pulling ahead of it) and the policy priority is to identify the emerging dominant channel and build the institutional capacity to accelerate it before the window for leapfrogging closes.")
)

cat(glue("
<div class='strategy-close'>
{closing_statement}
</div>
"))
```
